### 2017-04-24 articleBoston.tex
first commit

\documentclass[a4,11pt,twocolumn]{article}

\input{packageBoston}

\title{Creation of a corpus of realistic urban sound scenes with controlled acoustic properties}     %% \title est une macro, entre { } figure son premier argument
\author{
  GLOAGUEN, Jean-R\'emy\      \texttt{first1.last1@xxxxx.com}
  \and
  CAN, Arnaud\      \texttt{first2.last2@xxxxx.com}
  \and
  LAGRANGE, Mathieu\      \texttt{first1.last1@xxxxx.com}
  \and
  PETIOT, Jean-Fran\c cois\      \texttt{first1.last1@xxxxx.com}
}

\begin{document}

\maketitle

\section*{Abstract}
An approach is proposed to simulate a panel of urban sound mixtures from a study of recorded scenes. The aim is to succeed to compose sufficiently realistic scenes to be similar to recordings realized in cities. From a listening and recording phase, some parameters are extracted and can be used in the sound mixture simulator \textit{simScene}. The real scenes are then replicated and the realism of these are compared to the recorded scenes by a perceptive test realized on 50 people. The result prove that no particular differences can be made between the simulated and the real scenes.

\section*{Introduction}

The urban sound environment is the subject of many researches both on perspective \cite{jin_yong_soundwalk_2013} \cite{botteldooren_understanding_2011} and on physical aspects \cite{can_describing_2015} \cite{raimbault_ambient_2003}. These studies can be based on real sound environments either by acoustic listening through a \textit{soundwalk} \cite{adams_soundwalking_2008} or with recording made in city \cite{botteldooren_temporal_2006} or in a laboratory by using sound mixtures resulting from a simulation process \cite{lafay_new_2014}. In the first case, if the urban sound environment have an undeniable ecology validity, it doesn't make it possible to have a repeatable experience nor to propose a controlled framework where the presence and the level of the different sound sources can be adjusted. Then the use of a simulation tool to create specific urban sound environments is a the only way to control the presence of each class. It allows for instance to create urban sound environments by adjusting some parameters \cite{bruce_development_2009} or to test some classification or detection tools \cite{giannoulis_detection_2013} .

The main question is then to get scenes that are realistic enough to be similar to the recordings. Various methods have been proposed. The first approach would be to simulate completely a neighborhood taking into account the architecture of the buildings, the sound dynamic of the different sources present (car, voice, bird, bell \dots) and their propagation to a receptor, as proposed by \cite{cstb_simulation_2015}. If the first renderings are interesting, this method stays complicated to implement with a final rendering that is again too artificial. 
The second approach composes sound mixtures from isolated sound associated together. The idea is to consider the urban sound environment as the sum of acoustic events, i.e events with a sufficient sound level to be discernible, superposed to a sound background, i.e a sound constant on the sound mixture whose properties vary slowly during the scene. The difficulty is to create a representative database of isolated sounds with a sufficient quality to not deteriorate the sound quality and to put the acoustic events wisely to compose correct sound environments. The method proposed in \cite{misra_musical_2007} enables to resolve the first issue by extracting the acoustical event directly in real recordings and by manipulating them to re-use them in new mixtures. If the tool has a sufficient number of parameters to control the sound event, their method is limited by the extraction phase: the overlapping between the acoustics events deteriorates the sounds and then the final rendering. In more, the manipulation (duration, frequency range) can create artifacts that can be a limitation for a realistic use. In a similar way, the tool proposed by Davis and Bruce \cite{bruce_development_2009} enables to compose sound mixtures with an isolated sound database composed of recordings made specially for their application. The creation of the urban sound mixtures is made possible with the perceptual evaluation of a panel of listeners of the different sound environments. This method enables to determine which sounds are the most representative or the most linked to an urban sound environment from a perspective point of view.

However, as the objective is to get simulated scenes that are as realistic than recordings, this method can not be used here because if this method proposes a perspective validity it can not provide an ecological one. Thus, this study proposes to create urban sound mixtures from the listening of urban audio recordings and the estimation of some parameters extracted from them to use it in an audio simulation tool. The realism of the composed scene is then tested in a perspective test. The first part focuses on the study of a recordings made in Paris, then the second part deals with the simulation tool and the last part summarizes the result obtains in the perspective test.

\section{Study of realistic scenes}
Real urban noise recordings are listening in details to determine the composition of typical urban sound environment in terms of sound sources, presence and level. The recordings used have been made on the \nth{13} district of Paris (France) in frame of the GRAFIC project context \cite{aumond_sound_2016} during a soundwalk which was designed to cover different types of urban sound environments. The walk was 2.1 km long, consisting of 19 stop points of 3 to 5 minutes of listening each (see figure \ref{fig:soundwalk}). It was roamed on two days (03/23/2015 and 03/30/2015) twice a day (on the morning and on the afternoon) and in one direction (from West to East, WE) and then in the other (from East to West, EW). The acquisition system used was a ASA SENS and was carry on the backpack of the operator which was walking among the participants. In the end, 74 audio in a wav format and sampled at 44,1 kHz are available and serve as a base for the study. More details can be found on  \cite{aumond_modelling_2017}. \\

\begin{figure}[H]
\centering
\includegraphics[width=.5\textwidth]{./pictures/trajet_19pts.png}
\caption{Map of the soundwalk with the 19 stop points}
\label{fig:soundwalk}
\end{figure}

Each audio is annotated to determine the sound classes that can be present in an urban area, their recurrence and is classified among 4 possible sound environments (\textit{park}, \textit{quiet street} , \textit{noisy street}, \textit{very noisy street}) as proposed by \cite{rychtarikova_soundscape_2013} or \cite{can_describing_2015} (table~\ref{tab:classificationScene}).

\begin{table*}[h]
\centering

\begin{tabular}{|c|c|*{19}{l|}}
\hline
\multicolumn{1}{|l|}{\textbf{day}} & \textbf{journey}   & 1                        & 2                        & 3                        & 4                        & 5                        & 6                        & 7                        & 8                        & 9                        & 10                       & 11                       & 12                       & 13                       & 14                       & 15                       & 16                       & 17                       & 18                       & 19                       \\ \hline
\multirow{2}{*}{\textbf{1}} & \textbf{EW} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} \\ 
\cline{2-20} 
 & \textbf{WE} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} &                          & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} &  \\ 
 \hline
\multirow{2}{*}{\textbf{2}} & \textbf{EW} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} \\ 
\cline{2-20} 
 & \textbf{WE} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{5AB25A} \\ \hline
\end{tabular}

\vspace{0.5cm}

\begin{tabular}{|p{1cm}|l|p{0.001cm}|p{2cm}|l|p{0.001cm}|p{2cm}|l|p{0.001cm}|p{2.75cm}|l|p{0.001cm}|p{2cm}|l|}

\hhline{|-|-|~|-|-|~|-|-|~|-|-|~|-|-|}
park & {\cellcolor[HTML]{5AB25A}} & & quiet street & {\cellcolor[HTML]{FFCB2F}} & & noisy street & {\cellcolor[HTML]{F56B00}} & &  very noisy street & {\cellcolor[HTML]{9A0000}} & & unclassified & \\
\hhline{|-|-|~|-|-|~|-|-|~|-|-|~|-|-|}

\end{tabular}

\caption{Classification of the scenes according to the sound environment}
\label{tab:classificationScene}
\end{table*}

Most of the scene of the recordings belongs to the ambiance \textit{quiet street} and \textit{noisy street}. 2 scenes are not available : a noisy lorry sweeping the gutter is present throughout the \textit{1-WE-3} polluting the sound scene and the \textit{1-WE-19} record has been cut too quickly to be usable. From the annotation, each sound ambiance is characterized with the sound classes presents in it. The most common sound classes are resumed in the table \ref{tab:obsScene}.\\

\begin{table*}[!h]
\centering
\begin{tabular}{L{2.5cm} | C{2.5cm} | c | C{3cm} | C{2.5cm}}
\centering \textbf{Sound environment} & \textbf{Sound level (dB)} & \textbf{Background                                                       } &\textbf{Event} & \textbf{number events/min} \\ \hline
Park              & 69.0        & 
\begin{tabular}[c]{@{}l@{}}voice, \\ bird's whistles\end{tabular} & 
\begin{tabular}[c]{@{}l@{}}road traffic \\ voices\\ bird's whistle\\ street noise\\ foot step\end{tabular}                                         & 
\begin{tabular}[c]{@{}l@{}}0.5\\ 0.5\\ 0.5\\ 0.5\\ 0.3\end{tabular}                       \\\hline
Quiet street      & 70.2        & 
\begin{tabular}[c]{@{}l@{}}road traffic\\ bird\end{tabular}       & 
\begin{tabular}[c]{@{}l@{}}road traffic \\ voices\\ street noises\\ foot step\\ bird\\ \begin{tabular}[c]{@{}l@{}}construction \\ \hfill site noise \end{tabular}\\ door house\\ door car\end{tabular} & 
\begin{tabular}[c]{@{}l@{}}1.0\\ 0.7\\ 0.7\\ 0.5\\ 0.2\\ \begin{tabular}[c]{@{}l@{}} 0.1\\~\end{tabular}\\  0.2\\ 0.2\end{tabular}    
\\ \hline
Noisy street      & 73.5        & road traffic                                                      & \begin{tabular}[c]{@{}l@{}}traffic\\ foot step\\ voice\\ street noise\\ bell\\ bird\\ car horn\\ car's door\\ siren\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}9.0 \\ 0.5\\ 0.6\\ 0.4\\ 0.1\\ 0.2\\ 0.3\\ 0.2\\ 0.1\end{tabular} \\ \hline
Very noisy street & 76.0        & road traffic                                                      & \begin{tabular}[c]{@{}l@{}}traffic\\ voice\\ siren\\ car horn\\ bird\\ foot step\\ car's door\\ street noise\end{tabular}                          & \begin{tabular}[c]{@{}l@{}}40\\ 0.3\\ 0.2\\ 0.3\\ 0.2\\ 0.3\\ 0.2\\ 0.3\end{tabular}      
\end{tabular}
\caption{Sound level and classes presents depending on the sound environment}
\label{tab:obsScene}
\end{table*}

%\begin{table*}[t]
%\centering
%
%\begin{tabular}{p{3cm}p{2cm}p{10.5cm}}
%\hline
%\multicolumn{1}{c}{\textbf{Ambience}} &\textbf{\begin{tabular}[c]{@{}c@{}}level\\ (dB SPL)\end{tabular}} & \multicolumn{1}{c}{\textbf{Description}} \\ \hline
%park & \centering 69,0 & predominance of children's and adults' voices and bird whistles as sound events and backgrounds, the road traffic is low almost non-existent ($\approx$ 1 car/min) \\ 
%\hline
%quiet street & \centering 70,2 & The background noise is caused by a low road traffic with few car passages ($\approx$ 2 car/min), the birds are present and audible, the voice is composed of children's and adult's voices. The effect of reverberation on the street, due to low background noise, is present.\\ 
%\hline
%animated street & \centering 73,5 & traffic noise  is louder and the unique background, the car passages density is higher ($\approx$ 10 car/min), the human activity is more present and diversify (voice, step, suitcase, construction site noise \dots), the birds disappear. \\ 
%\hline
%very noisy street & \centering 76,0 & the background noise very high composed of quasi-constant traffic flow ($\sup$ 40 car/min), the human activity are less presents in a favor of voice hubbub.\\  
%\hline   
%
%\end{tabular}
%\caption{Qualitative observations on the urban sound environment}
%\label{tab:obsScene}
%\end{table*}

The road traffic, the voice and the bird components are the most linked to the sound atmosphere: with exclusively road traffic, the sound environment is more noisy than in the \textit{park} and \textit{quiet street} where the voice and the bird's whistles are predominant. Furthermore, multiple sounds classes can be heard whatever the urban sound environments : dog barks, church bell ringing, car horn, sirens \dots with a density which can be very low (< 0.1/min). Finally, a lot of brief mechanics sounds with unknown origin can be listened in almost all the audio. Consequently, all theses indefinite sounds are classed in one unique class sound, \textit{street noise}. From theses information (density, level, sound classes present in each ambience), it is now possible to compose urban sound mixtures.

\section{Creation of realistic simulated urban scenes}
\subsection{Presentation of the web-simulator \textit{simScene}}
\textit{simScene} \cite{rossignol_simscene:_2015} is a web-simulator that creates sound mixtures in a wav format by superposing audio which comes from an isolated sound database composed of two categories: the brief sounds (from 1 to 20 seconds)
that can be salient are belong to the \textit{event} category and all the sound that are long and whose acoustic properties do not vary are put in the \textit{background} category. In each category, the sounds are then regrouping in some sound classes (bird, car, foot steps \dots). The software enables to control for each sound classes some parameters as the number of events of each class that appears in the mixture, the elapsed time between each sample of a same class or the existence of a fade in and a fade out \dots. Each parameters are completed with a standard deviation bringing some random behavior between the scenes. \\

\begin{figure}[h]
\centering
\includegraphics[width = \linewidth]{./pictures/spectrogramme_abstract_sceneSimpleKlaxonFixe_EN.pdf}
\caption{Example of a scene composed of three sound classes}
\label{fig:exampleSimScene}
\end{figure}

The sound mixtures can be create following two modes: the \textit{abstract} mode allows to create mixtures from the specified parameters while \textit{replicate} mode reproduces an existing scene from its annotation file which resumes the sound classes corresponding, their time of apparition and disappearance of every events. With the global sound mixture, an audio can also be created for each sound class present in the scene which allows to know their exact contribution in the scene. In parallel, \textit{simScene} generates 3 image files representing the temporal evolution, the spectrogram and a piano Roll representation to visualize each sound classes distribution and also two text files summarizing the temporal location of each sample and all the parameter used. \\

\subsection{Creation of the urban sound database}
Based on the annotations of the real scenes, a sound database is built with sounds found online on the \textit{freesound} project and with the urbanSound8k database\cite{salamon_dataset_nodate}. This database consists in more than 8000 files with a 4 seconds or less duration found on the web site freesound.org too, and classified in 10 sound classes : ventilation, car horn, children playing, dog barking, bell, engine idling, gunshot, jackhammer, siren and street music. All the audio have been sorted and those who presented the best signal noise ratio and are present in the sound classes identified in the real records have been selected.\\

In addition, as road traffic is a prime audio source in an urban sound environment, it seemed interesting to have, in the database, a part composed of well-recorded car audio. As a result, passages of 4 different cars (Renault Scenic, Clio, Megane and Dacia Sandero) were recorded on the Ifsttar-Nantes runway at different speeds and gear ratios for steady, acceleration and braking phases (table~\ref{tab:resum_car_audio}). In all, 103 car passages have been recorded. \\

\begin{table*}[!htb]
    \begin{minipage}{.5\linewidth}
      \centering
        \begin{tabular}{|c|c|lllll}
\cline{2-7}
\multicolumn{1}{l|}{} & \multicolumn{1}{l|}{\textbf{Trans.}} & \multicolumn{1}{l|}{\textbf{1}} & \multicolumn{1}{l|}{\textbf{2}} & \multicolumn{1}{l|}{\textbf{3}} & \multicolumn{1}{l|}{\textbf{4}} & \multicolumn{1}{l|}{\textbf{5}} \\ \hline
\multicolumn{1}{|c|}{\multirow{8}{*}{\begin{tabular}[c]{@{}c@{}}\textbf{stabilized}\\   \textbf{speed}\\   \textbf{(km/h)}\end{tabular}}} & \textbf{20} & \multicolumn{1}{l|}{$\times$} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} \\ \cline{2-7} 
\multicolumn{1}{|c|}{} & \textbf{30} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{$\times$} & \multicolumn{1}{l|}{$\times$} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} \\ \cline{2-7} 
\multicolumn{1}{|c|}{} & \textbf{40} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{$\times$} & \multicolumn{1}{l|}{$\times$} & \multicolumn{1}{l|}{$\times$} & \multicolumn{1}{l|}{} \\ \cline{2-7} 
\multicolumn{1}{|c|}{} & \textbf{50} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{$\times$} & \multicolumn{1}{l|}{$\times$} & \multicolumn{1}{l|}{} \\ \cline{2-7} 
\multicolumn{1}{|c|}{} & \textbf{60} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{$\times$} & \multicolumn{1}{l|}{$\times$} \\ \cline{2-7} 
\multicolumn{1}{|c|}{} & \textbf{70} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{$\times$} & \multicolumn{1}{l|}{$\times$} \\ \cline{2-7} 
\multicolumn{1}{|c|}{} & \textbf{80} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{$\times$} \\ \cline{2-7} 
\multicolumn{1}{|c|}{} & \textbf{90} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{$\times$} \\ \hline
\end{tabular}
    \end{minipage}
    \begin{minipage}{.5\linewidth}
      \centering
\begin{tabular}{|c|c||c|c|}
\hline
\multicolumn{2}{|c||}{\textbf{braking}} & \multicolumn{2}{|c|}{\textbf{acceleration}} \\ \hline
\multicolumn{1}{|c|}{\begin{tabular}[c]{@{}c@{}}\textbf{speed}\\  \textbf{(km/h)}\end{tabular}} & \multicolumn{1}{c||}{\textbf{trans.}} & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}\textbf{speed}\\  \textbf{(km/h)}\end{tabular}} & \multicolumn{1}{c|}{\textbf{trans.}} \\ \hline
50 → 0 & 3 → 2 & 0 → 30 & 1 → 2 \\ \hline
40 → 0 & 2 → 2 & 0 → 40 & 1 → 2 \\ \hline
50 → 30 & 3 → 2 & 20 → 40 & 1 → 3 \\ \hline
60 → 40 & 4 → 3 & 30 → 50 & 2 → 3 \\ \hline
70 → 50 & 4 → 3 & 40 → 60 & 3 → 4 \\ \hline
80 → 50 & \begin{tabular}[c]{@{}l@{}}4 or 5\\ → 3\end{tabular} & 50 → 70 & \begin{tabular}[c]{@{}l@{}}3 → \\ 4 or 5\end{tabular} \\ \hline
\end{tabular}
    \end{minipage}%
    \caption{Measurement set on runway with vehicle passages with stabilized speed (left) and in acceleration and braking phase(right)} 
    \label{tab:resum_car_audio}
\end{table*}

A median filter \cite{fitzgerald_harmonic/percussive_2010} has been applied to filter out bird's whistles from the recordings in the frequency range $\left[ 2500-6500 \right]$ Hz. It   consists in taking a window around a center point and to attribute to it the median value of this window. This method allow to attenuate the brief variations both in the harmonic and the temporal plan. The size of the filter window is chosen with a 5 points width of and a 9 points height so that the effect of the filter is sufficient without deteriorating the quality of the audio (see figure \ref{fig:filtre_car}).

\begin{figure}[hbtp]
\centering
\includegraphics[width=\linewidth]{./pictures/filtrageMedian_VL1_R3_40_EN.pdf}
\caption{Spectrogram of a recorded passage of the Renault Megane car at 40 km/h with the \nth{3} gear ratio ($N_{w} = 2^{12}$ with 50 $\%$ overlapping, $N_{fft} = 2^{12}$, $sr = 44,1$ kHz). On the left, the the original recording, on the right, the filtered audio with the filtered part framed.}
\label{fig:filtre_car}
\end{figure}

Finally, the built database is composed of 245 sound events (bell, whistle bird, sweeping broom, car horn, car, hammer and drill, coughing, dog barking, car and house door slamming, plane, siren, foot step, thunder, street noise, suitcase rolling, train and tramway passing, truck and voice) and of 154 background sounds (birds, construction site, crowd, park, rain, schoolyard, traffic, ventilation, wind). With this built-up database, the 74 real sound scenes are replicated. To test whether the replicated scenes are sufficiently realistic, several of them are evaluated by a panel of listeners.

\section{Perceptual test}
\subsection{Creation of the test}
A perspective test is created where a panel of listeners evaluates the realism of a mix of real and replicated scenes on a 7-point scale (1 is \textit{not realistic at all}, 7 is \textit{very realistic}). Since it is not possible for each panelist to evaluate all the real and replicated scenes, in order to limit the duration of the test and to keep the judge's concentration constant, each listens a panel of 20 sound scenes composed of 10 real sounds scenes and 10 replicated. Furthermore, for the auditory comfort of the panelist, all the scenes are normalized to the same sound level, chosen at 65 dB.\\

First, the listening plan is elaborated following a Balanced Incomplete Block Design \cite{dagnelie_principes_2003} where $J$ participants realized $K$ evaluations from $B$ sound sources, each of them are tested $R$ times and each couple of sound sources are presented $\lambda$ times. 3 rules govern these parameters : 
\begin{subequations}\label{eq:BIE_cond}
\begin{align}
J &\geq B, \label{eq:BIE_cond1}\\
JK &= BR, \label{eq:BIE_cond2}\\
\lambda &= R\frac{K-1}{J-1}. \label{eq:BIE_cond3}
\end{align}
\end{subequations}

with $\left[J, B, K, R, \lambda\right] \in \mathbb{N}$. A reasonable number of listeners is chosen at $J = 50$ and the total numberof sound sources $B$ to be tested is set at 40. It consists of 20 30-seconds audios from the recorded scene and the same 30 seconds from the replicated scenes. Among the 20 scenes, there are 5 scenes which belong to the ambience \textit{Park}, 6 from \textit{Quiet street}, 4 from \textit{noisy street} and 5 from \textit{very noisy street} chosen randomly among the 74 scenes. As (\ref{eq:BIE_cond3}) is not satisfied, the listening plan is finally elaborated following an optimal plan \cite{pages_blocs_2007}. It consists, with the parameters $J$, $K$ and $R$ fixed, to determine the optimal plan that would be the most balanced (figure \ref{fig:repartition}).\\

\begin{figure}[ht]
\centering
\includegraphics[width = \linewidth]{./pictures/repartition-real-simulated.pdf}
\caption{Distribution between the real and simulated scenes for each judge. The cumulative sum is equal to the number of tested elements $K$.}
\label{fig:repartition}
\end{figure}

A web-page\footnote{http://soundthings.org/research/xpRealism} was uploaded on the 8 february 2007 and the number of 50 compiled test has been reached 12 days later. During the test, the participant had the possibility to listen each scene as many time as wanted before evaluating it without being able to change his judgment afterwards. He could also leave a comment on each audio to explain his note. Finally, his age, gender and experience on listening to urban sound mixtures were asked at the end of the test. In the end, the panel of 50 listeners was made of 31 males and 18 females (one not documented) with an average age of 36 ($\pm$ 12) years old. $62\%$ of the participants declared they had no experience in the listening of urban sound mixtures.\\

%\begin{figure}[ht]
%\centering
%\includegraphics[width=\linewidth]{../../../Pictures/test_perceptif/testPerceptif_panel.pdf} 
%\caption{Description of the panel information}
%\label{fig:panel}
%\end{figure}

\subsection{Results}

An analyze of variance (ANOVA) is performed to see if both type of scenes can be discernible. 2 hypothesis are then possible : $H_0$, the real and replicates scenes can not be discernible, and $H_1$, both scenes are perceived with differences. For that, an threshold significance $\alpha=0,05$ is compared to the \textit{p-value} calculated by the ANOVA. This value represents probability to have an extreme value or more than the actual data observed considering the $H_0$ hypothesis. Two situations can happen : 
\begin{itemize}
\item $\alpha > p-value$, the $H_1$ hypothesis is validated and $H_0$ is rejected, 
\item $\alpha < p-value$, the $H_0$ is not rejected.
\end{itemize}

The data are represented by a box-and-whiskers plot according their type. 
\begin{figure}[hbtp]
\centering
\includegraphics[width=\linewidth]{./pictures/testPerceptif_boxplotType_EN.pdf}
\caption{Box-and-whiskers plot according the type of scene}
\label{fig:boxplot_real_simul}
\end{figure}

The distribution of the note are extremely similar. The $p-value$ is calculated at $0.18 > \alpha$ which means that the two distribution cannot be distinguished. So the simulated scenes are perceived in a similar way than the recorded scenes. The mean for the simulated scene are even superior to the real ($m_{simul.} = 5.1 (\pm 1.6$), $ m_{real} = 4.9 (\pm 1.6$)). \\

Secondly, an ANOVA is performed to determine if the experience in the listening of urban sound mixtures is an influential factor to distinguish real and replicated scenes. It is then a 2 dimensions ANOVA which is used. The test allow to determine  the influence of each dimension separately and their interaction. In this case, it means that the comportment of a dimension is depending on those of the other dimension. The $p-value$ associated is submitted on the two-hypotheses test (table~\ref{tab:p_value_type_exp}).\\

\begin{table}[h]
\centering
\begin{tabular}{cccc}
\cline{2-4}
          & type & experience & interaction \\
\hline
$p-value$ & 0.50 & 0.05 & 0.17 \\
\hline    
\end{tabular}
\caption{$p-value$ for the ANOVA at two dimensions with the interaction effect}
\label{tab:p_value_type_exp}
\end{table}

The dimensions and their interaction are not significantly influential since their $ p-value $ is equal to or greater than the $ \alpha$ significance threshold. This means that even listeners trained to hear urban sound environments do not dissociate the real scenes from the simulated ones.\\

A last ANOVA is performed with 2 dimensions to see if the scenes according to the \textit{type} and their \textit{sound environment} are evaluated differently (figure~\ref{fig:boxplot_type_ambience} and table \ref{tab:p_value_type_ambience}).

\begin{figure}[h]
\centering
\includegraphics[width=\linewidth]{./pictures/testPerceptif_boxplotAmbianceCOLOR_EN.pdf}
\caption{Box-and-whiskers plot according the type and the sound environment}
\label{fig:boxplot_type_ambience}
\end{figure}

\begin{table}[h]
\centering
\begin{tabular}{cccc}
\cline{2-4}
          & type   & sound env. & interaction \\
\hline
$p-value$ & $0.13$ & $0.80 \times 10^{-3}$     & $2.20 \times 10^{-3}$ \\
\hline    
\end{tabular}
\caption{$p-value$ for the anova at two dimensions with the interaction effect}
\label{tab:p_value_type_ambience}
\end{table}

For the \textit{type} dimension, the effect is still not significance ($p-value > \alpha$) contrary to the \textit{sound environment} dimension. Thus, the realism of the listened scenes seems most influenced by the kind of sound environment than the type of sound scenes. As the mean of the \textit{noisy} and \textit{very noisy} atmospheres are higher, it seems that the presence of cars are a preponderant element to evaluate the realism of an audio mixture. To the opposite, for the \textit{quiet} and \textit{park} atmosphere, the evaluation is more disperse. From the comments let by the panelists, in the real scenes, the presence of too loud birds and some street noises with unknown origins have degraded the evaluation. Whereas in the simulated scenes, it is mainly the sound class \textit{foot step} and a sound background composed of birds that made the realism of the scenes decrease. \\

\section*{Conclusion}
Realistic urban sound mixtures have been composed from the study of urban recordings. A work of annotation has allowed to extract, for different acoustic atmosphere, some useful information as the sound classes presents, the traffic flow rates, the sound level. Form these observation, a urban sound database have been created including car passages recordings. A perceptual test has been set up to prove the realism of the sound mixtures. This test revealed that the replicated scenes are mistaken with the urban recordings. Furthermore, the results shows that it is more the sound environment which impact the realism than the experience of the listener. In the \textit{park} and \textit{quiet street} environment, the evaluation are more disperses and the mean note lower than the \textit{noisy} and \textit{very noisy} atmosphere. Theses differences come from mainly from the sound level of some events which are too loud to be realistic enough but that can be improved. From theses issues, it is possible to improve again the realism  of the replicated scenes by taking these comments account.

\section*{Acknowledge}
We would like to thank Pierre Aumond and Catherine Lavandier from the University of Cergy-Pontoise for transmitting us the data of the \textit{Grafic} project.

\bibliographystyle{unsrt}
\bibliography{../bibliographie}

\end{document}

### 2017-04-24 packageBoston.tex
first commit

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}

%%%%%% MISE EN PAGES %%%%%%
\usepackage[left=2cm, right=2cm, top=2cm]{geometry}

\setcounter{tocdepth}{3}     % Dans la table des matieres
\setcounter{secnumdepth}{3}  % Avec un numero.

%%%%%% SYMBOLES %%%%%
\usepackage{tipa}	% pour avoir l'accent concave
\usepackage{nth} 	% pour avoir first, second ... : \nth{num}
%%%%%% EQUATION %%%%%%
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{fancybox}
\usepackage{xfrac}	% fraction de type "1/4"
\usepackage{cases}	% système équation
\usepackage[overload]{empheq}
\usepackage{bm}		% pour mettre en gras .
\usepackage{units} 	% x/y barre latérale pour les fractions

%%%%%% FIGURE %%%%%%
\usepackage{graphicx}	% insérer des graphiques
\usepackage{subfigure}	% utiliser subfigure
\usepackage{float}	% utiliser H dans les figures

%%%%%% TABLEAUX %%%%%%
\usepackage{array,multirow,makecell}
\addto\captionsfrench{\def\tablename{\textsc{Tableau}}}% pour avoir TABLEAU et pas TABLE dans les légendes des tableaux
\usepackage[table,xcdraw]{xcolor} % pour avoir des lignes colorées dans les tableau
\usepackage{slashbox} % pour les \backslashbox
%\usepackage{subcaption}
\usepackage{hhline}	% pour les lignes horizontales

\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

%%%%%%%%%%%%%%%%%%%%%
\linespread{1.2}	% interligne

### 2017-04-24 articleBoston.tex
bibliography correection
title and abstract

@@
 
-\title{Creation of a corpus of realistic urban sound scenes with controlled acoustic properties}     %% \title est une macro, entre { } figure son premier argument
+\title{Design of a urban sound scenes corpuses with controlled acoustic properties}     %% \title est une macro, entre { } figure son premier argument
 \author{
@@
 \section*{Abstract}
-An approach is proposed to simulate a panel of urban sound mixtures from a study of recorded scenes.
-The aim is to succeed to compose sufficiently realistic scenes to be similar to recordings realized in cities.
-From a listening and recording phase, some parameters are extracted and can be used in the sound mixture simulator \textit{simScene}.
-The real scenes are then replicated and the realism of these are compared to the recorded scenes by a perceptive test realized on 50 people.
-The result prove that no particular differences can be made between the simulated and the real scenes.
+
+Monitoring the acoustic urban environment is beneficial for improving the life of citizens.
+To do so, the research community in acoustics can befenit from corpora where the acoustic contribution of sources of disturbance such as road traffic is precisely known.
+
+In this paper, a set of urban sound mixtures are simulated using an open source tool \textit{simScene}.
+Their caracteristics (type of sources, density of events, etc.) are infered from the listening of reference audio scenes recorded in Paris, France.
+The main advantage of the simulated versions is that several quantities, such as the acoustic level of each source is precisely known, a useful feature for evaluated automated analysis tools.
+
+In order to quantifiy the level of realism achieved by the use of the simulation tool a perceptive test is conducted with 50 subjects.
+The result show that no significative differences can be made between the simulated and the recorded scenes.
 
@@
 \bibliographystyle{unsrt}
-\bibliography{../bibliographie}
+\bibliography{bibliography}
 

### 2017-04-24 articleBoston.tex
fix intro

@@
 
-The urban sound environment is the subject of many researches both on perspective \cite{jin_yong_soundwalk_2013} \cite{botteldooren_understanding_2011} and on physical aspects \cite{can_describing_2015} \cite{raimbault_ambient_2003}.
+The urban sound environment is studied both from the perception \cite{jin_yong_soundwalk_2013} \cite{botteldooren_understanding_2011} and from the physical point of view \cite{can_describing_2015} \cite{raimbault_ambient_2003}.
 These studies can be based on real sound environments either by acoustic listening through a \textit{soundwalk} \cite{adams_soundwalking_2008} or with recording made in city \cite{botteldooren_temporal_2006} or in a laboratory by using sound mixtures resulting from a simulation process \cite{lafay_new_2014}.
 In the first case, if the urban sound environment have an undeniable ecology validity, it doesn't make it possible to have a repeatable experience nor to propose a controlled framework where the presence and the level of the different sound sources can be adjusted.
-Then the use of a simulation tool to create specific urban sound environments is a the only way to control the presence of each class.
-It allows for instance to create urban sound environments by adjusting some parameters \cite{bruce_development_2009} or to test some classification or detection tools \cite{giannoulis_detection_2013} .
+Then the use of a simulation tool to create specific urban sound environments is a useful way to control the degree of contribution of each class of acoustical sources.
+It allows us to create urban sound environments by adjusting some parameters \cite{bruce_development_2009} or to evaluate the performance of some classification or detection tools \cite{giannoulis_detection_2013}.
 
-The main question is then to get scenes that are realistic enough to be similar to the recordings.
-Various methods have been proposed.
+A critical question is then to obtain scenes that are realistic enough, thus to be considered as similar to some reference recordings.
+To do so, various methods have been proposed.
 The first approach would be to simulate completely a neighborhood taking into account the architecture of the buildings, the sound dynamic of the different sources present (car, voice, bird, bell \dots) and their propagation to a receptor, as proposed by \cite{cstb_simulation_2015}.
-If the first renderings are interesting, this method stays complicated to implement with a final rendering that is again too artificial.
-The second approach composes sound mixtures from isolated sound associated together.
-The idea is to consider the urban sound environment as the sum of acoustic events, i.e events with a sufficient sound level to be discernible, superposed to a sound background, i.e a sound constant on the sound mixture whose properties vary slowly during the scene.
-The difficulty is to create a representative database of isolated sounds with a sufficient quality to not deteriorate the sound quality and to put the acoustic events wisely to compose correct sound environments.
+If the first renderings are interesting, this method remains complex to implement.
+A more straitghtforward approach composes sound mixtures from isolated sound mixed together.
+The idea is to consider the urban sound environment as the sum of acoustic events, i.e.
+events with a sufficient sound level to be discernible, superposed to a sound background, i.e a sound constant on the sound mixture whose properties vary slowly during the scene.
+The difficulty here is to have a representative database of isolated sounds with a sufficient quality not to compomise the resulting sound quality and to sequence the acoustic events wisely.
+
 The method proposed in \cite{misra_musical_2007} enables to resolve the first issue by extracting the acoustical event directly in real recordings and by manipulating them to re-use them in new mixtures.
@@
 
-However, as the objective is to get simulated scenes that are as realistic than recordings, this method can not be used here because if this method proposes a perspective validity it can not provide an ecological one.
-Thus, this study proposes to create urban sound mixtures from the listening of urban audio recordings and the estimation of some parameters extracted from them to use it in an audio simulation tool.
-The realism of the composed scene is then tested in a perspective test.
-The first part focuses on the study of a recordings made in Paris, then the second part deals with the simulation tool and the last part summarizes the result obtains in the perspective test.
+However, as our objective here is to get simulated scenes that are as realistic as possible, this method can not be considered because if this method proposes a perspective validity it can not provide an ecological one.
+Thus, this study proposes to create urban sound mixtures from the listening of urban audio recordings and the estimation of some high level sequencing parameters extracted from the recordings to tune an audio simulation tool.
+The realism of the simulated scenes is then evaluated using a perspective test.
 
-\section{Study of realistic scenes}
+The remaining of the paper is organized as follows, Section \ref{sec:study} focuses on the study of recordings made in Paris.
+Section \ref{sec:simulation} deals with the simulation tool and Section \ref{sec:test} summarizes the result obtained in the perspective test.
+
+\section{Study of realistic scenes} \label{sec:study}
 Real urban noise recordings are listening in details to determine the composition of typical urban sound environment in terms of sound sources, presence and level.
@@
 
-\section{Creation of realistic simulated urban scenes}
+\section{Simulation of realistic simulated urban scenes} \label{sec:simulation}
 \subsection{Presentation of the web-simulator \textit{simScene}}
@@
 
-\section{Perceptual test}
+\section{Perceptual test} \label{sec:test}
 \subsection{Creation of the test}
@@
 
-\section*{Acknowledge}
+\section*{Acknowledgements}
 We would like to thank Pierre Aumond and Catherine Lavandier from the University of Cergy-Pontoise for transmitting us the data of the \textit{Grafic} project.

### 2017-04-24 articleBoston.tex
fix sec 1

@@
 \section{Study of realistic scenes} \label{sec:study}
-Real urban noise recordings are listening in details to determine the composition of typical urban sound environment in terms of sound sources, presence and level.
-The recordings used have been made on the \nth{13} district of Paris (France) in frame of the GRAFIC project context \cite{aumond_sound_2016} during a soundwalk which was designed to cover different types of urban sound environments.
+Real urban noise recordings are listened in details to determine the composition of typical urban sound environment in terms of sound sources, presence and level.
+The recordings considered in this study have been made on the \nth{13} district of Paris (France) as part of the GRAFIC project \cite{aumond_sound_2016} during a soundwalk which was designed to cover different types of urban sound environments.
 The walk was 2.1 km long, consisting of 19 stop points of 3 to 5 minutes of listening each (see figure \ref{fig:soundwalk}).
 It was roamed on two days (03/23/2015 and 03/30/2015) twice a day (on the morning and on the afternoon) and in one direction (from West to East, WE) and then in the other (from East to West, EW).
-The acquisition system used was a ASA SENS and was carry on the backpack of the operator which was walking among the participants.
-In the end, 74 audio in a wav format and sampled at 44,1 kHz are available and serve as a base for the study.
+The acquisition system used was a ASA SENS and was carried on the backpack of the operator which was walking with the participants.
+In the end, 74 audio pulse code modulation recordings sampled at 44,1 kHz are available and serve as a base for the study.
 More details can be found on  \cite{aumond_modelling_2017}.
@@
 
-Each audio is annotated to determine the sound classes that can be present in an urban area, their recurrence and is classified among 4 possible sound environments (\textit{park}, \textit{quiet street} , \textit{noisy street}, \textit{very noisy street}) as proposed by \cite{rychtarikova_soundscape_2013} or \cite{can_describing_2015} (table~\ref{tab:classificationScene}).
+Each audio is fully annotated in terms of sound classes that are be present in a urban area, their recurrence and is labeled using the following types of sound environments (\textit{park}, \textit{quiet street} , \textit{noisy street}, \textit{very noisy street}) as proposed by \cite{rychtarikova_soundscape_2013} or \cite{can_describing_2015} (table~\ref{tab:classificationScene}).
 
@@
 
-Most of the scene of the recordings belongs to the ambiance \textit{quiet street} and \textit{noisy street}.
-2 scenes are not available : a noisy lorry sweeping the gutter is present throughout the \textit{1-WE-3} polluting the sound scene and the \textit{1-WE-19} record has been cut too quickly to be usable.
+Most of the scene of the recordings belong to \textit{quiet street} and \textit{noisy street}.
+2 scenes are not available : a noisy cleaning truck is present throughout the \textit{1-WE-3} heavily polluting the sound scene and the \textit{1-WE-19} record is too short to be considered in this study.
 From the annotation, each sound ambiance is characterized with the sound classes presents in it.
-The most common sound classes are resumed in the table \ref{tab:obsScene}.\\
+The most common sound classes are displayed in Table \ref{tab:obsScene}.\\
 
@@
 Finally, a lot of brief mechanics sounds with unknown origin can be listened in almost all the audio.
-Consequently, all theses indefinite sounds are classed in one unique class sound, \textit{street noise}.
+Consequently, all these indefinite sounds are annotated in one unique class sound, \textit{street noise}.
 From theses information (density, level, sound classes present in each ambience), it is now possible to compose urban sound mixtures.
@@
 \subsection{Presentation of the web-simulator \textit{simScene}}
-\textit{simScene} \cite{rossignol_simscene:_2015} is a web-simulator that creates sound mixtures in a wav format by superposing audio which comes from an isolated sound database composed of two categories: the brief sounds (from 1 to 20 seconds)
-that can be salient are belong to the \textit{event} category and all the sound that are long and whose acoustic properties do not vary are put in the \textit{background} category.
-In each category, the sounds are then regrouping in some sound classes (bird, car, foot steps \dots).
-The software enables to control for each sound classes some parameters as the number of events of each class that appears in the mixture, the elapsed time between each sample of a same class or the existence of a fade in and a fade out \dots.
-Each parameters are completed with a standard deviation bringing some random behavior between the scenes.
+\textit{simScene} \cite{rossignol_simscene:_2015} is a simulator that creates sound mixtures in a wav format by superposing audio which comes from an isolated sound database composed of two categories: the brief sounds (from 1 to 20 seconds)
+that are considered as salient belong to the \textit{event} category and all the sound that are of long duration and whose acoustic properties do not vary with respect to time belong to the \textit{background} category.
+In each category, the sounds are then grouped in sound classes (bird, car, foot steps \dots).
+The software enables the user to control for each sound classes some parameters as the number of events of each class that appears in the mixture, the elapsed time between each sample of a same class or the existence of a fade in and a fade out \dots.
+Each parameters are completed with a standard deviation optionally bringing some random behavior between the scenes.
 \\
@@
 
-The sound mixtures can be create following two modes: the \textit{abstract} mode allows to create mixtures from the specified parameters while \textit{replicate} mode reproduces an existing scene from its annotation file which resumes the sound classes corresponding, their time of apparition and disappearance of every events.
+The sound mixtures can be created following two modes: the \textit{abstract} mode allows us to create mixtures from the specified parameters while the \textit{replicate} mode reproduces an existing scene by following as much as possible its annotation file.
 With the global sound mixture, an audio can also be created for each sound class present in the scene which allows to know their exact contribution in the scene.
-In parallel, \textit{simScene} generates 3 image files representing the temporal evolution, the spectrogram and a piano Roll representation to visualize each sound classes distribution and also two text files summarizing the temporal location of each sample and all the parameter used.
+%In parallel, \textit{simScene} generates 3 image files representing the temporal evolution, the spectrogram and a piano Roll representation to visualize each sound classes distribution and also two text files summarizing the temporal location of each sample and all the parameter used.
 \\
@@
 \subsection{Creation of the urban sound database}
-Based on the annotations of the real scenes, a sound database is built with sounds found online on the \textit{freesound} project and with the urbanSound8k database\cite{salamon_dataset_nodate}.
+Based on the annotations of the real scenes, a sound database is built using the \textit{replicate} mode with sounds found online on the \textit{freesound} project and with the urbanSound8k database\cite{salamon_dataset_nodate}.
 This database consists in more than 8000 files with a 4 seconds or less duration found on the web site freesound.org too, and classified in 10 sound classes : ventilation, car horn, children playing, dog barking, bell, engine idling, gunshot, jackhammer, siren and street music.
-All the audio have been sorted and those who presented the best signal noise ratio and are present in the sound classes identified in the real records have been selected.\\
+% ML Pas clair, est-ce utile ?
+All the audio have been sorted and those who presented the best signal noise ratio are selected in the sound classes identified in the real records have been selected.\\
 
@@
 
-Finally, the built database is composed of 245 sound events (bell, whistle bird, sweeping broom, car horn, car, hammer and drill, coughing, dog barking, car and house door slamming, plane, siren, foot step, thunder, street noise, suitcase rolling, train and tramway passing, truck and voice) and of 154 background sounds (birds, construction site, crowd, park, rain, schoolyard, traffic, ventilation, wind).
+Finally, the resulting database is composed of 245 sound events (bell, whistle bird, sweeping broom, car horn, car, hammer and drill, coughing, dog barking, car and house door slamming, plane, siren, foot step, thunder, street noise, suitcase rolling, train and tramway passing, truck and voice) and of 154 background sounds (birds, construction site, crowd, park, rain, schoolyard, traffic, ventilation, wind).
 With this built-up database, the 74 real sound scenes are replicated.

### 2017-04-24 articleBoston.tex
done with proof reading

@@
 
-\title{Design of a urban sound scenes corpuses with controlled acoustic properties}     %% \title est une macro, entre { } figure son premier argument
+\title{Design and Perceptual Evaluation of a urban sound scenes corpus with controlled acoustic properties}     %% \title est une macro, entre { } figure son premier argument
 \author{
@@
 \subsection{Creation of the test}
-A perspective test is created where a panel of listeners evaluates the realism of a mix of real and replicated scenes on a 7-point scale (1 is \textit{not realistic at all}, 7 is \textit{very realistic}).
-Since it is not possible for each panelist to evaluate all the real and replicated scenes, in order to limit the duration of the test and to keep the judge's concentration constant, each listens a panel of 20 sound scenes composed of 10 real sounds scenes and 10 replicated.
-Furthermore, for the auditory comfort of the panelist, all the scenes are normalized to the same sound level, chosen at 65 dB.\\
+A perceptive test is conduected with a panel of listeners that are asked to evaluate the level of realism on a 7-point scale (1 is \textit{not realistic at all}, 7 is \textit{very realistic}).
+Evaluated scenes are  recorded and replicated ones.
+In order to limit the duration of the test and to preserve the concentration of the subjects, each subject listens a panel of 20 sound scenes composed of 10 real sounds scenes and 10 replicated among the 40 scenes available.
+All the scenes are normalized to the same sound level, chosen at 65 dB.\\
 
-First, the listening plan is elaborated following a Balanced Incomplete Block Design \cite{dagnelie_principes_2003} where $J$ participants realized $K$ evaluations from $B$ sound sources, each of them are tested $R$ times and each couple of sound sources are presented $\lambda$ times.
+The listening plan is elaborated following a Balanced Incomplete Block Design \cite{dagnelie_principes_2003} where $J$ participants realized $K$ evaluations from $B$ sound sources, each of them are tested $R$ times and each couple of sound sources are presented $\lambda$ times.
 3 rules govern these parameters :
@@
 with $\left[J, B, K, R, \lambda\right] \in \mathbb{N}$.
-A reasonable number of listeners is chosen at $J = 50$ and the total numberof sound sources $B$ to be tested is set at 40.
-It consists of 20 30-seconds audios from the recorded scene and the same 30 seconds from the replicated scenes.
+A reasonable number of listeners is chosen at $J = 50$ and the total number of sound sources $B$ to be tested is set at 40.
+It consists of 20 30-seconds audios of the recorded scenes and the same 30 seconds from the replicated scenes.
 Among the 20 scenes, there are 5 scenes which belong to the ambience \textit{Park}, 6 from \textit{Quiet street}, 4 from \textit{noisy street} and 5 from \textit{very noisy street} chosen randomly among the 74 scenes.
@@
 
-A web-page\footnote{http://soundthings.org/research/xpRealism} was uploaded on the 8 february 2007 and the number of 50 compiled test has been reached 12 days later.
+A web-page\footnote{\url{http://soundthings.org/research/xpRealism}} was uploaded on the 8 february 2007 and the number of 50 participants has been reached 12 days later.
 During the test, the participant had the possibility to listen each scene as many time as wanted before evaluating it without being able to change his judgment afterwards.
@@
 Finally, his age, gender and experience on listening to urban sound mixtures were asked at the end of the test.
-In the end, the panel of 50 listeners was made of 31 males and 18 females (one not documented) with an average age of 36 ($\pm$ 12) years old.
+The panel of 50 listeners was made of 31 males and 18 females (one not documented) with an average age of 36 ($\pm$ 12) years old.
 $62\%$ of the participants declared they had no experience in the listening of urban sound mixtures.\\
@@
 
-An analyze of variance (ANOVA) is performed to see if both type of scenes can be discernible.
-2 hypothesis are then possible : $H_0$, the real and replicates scenes can not be discernible, and $H_1$, both scenes are perceived with differences.
+An analysis of variance (ANOVA) is performed to check if there is a significant difference of realism between the recorded and the simulated scenes.
+2 hypothesis are evaluated: $H_0$, the recorded and simulated scenes are not different in terms of realism, and $H_1$, both scenes are perceived with differences.
 For that, an threshold significance $\alpha=0,05$ is compared to the \textit{p-value} calculated by the ANOVA.
-This value represents probability to have an extreme value or more than the actual data observed considering the $H_0$ hypothesis.
+This value represents the probability to have an extreme value or more than the actual data observed considering the $H_0$ hypothesis.
 Two situations can happen :
@@
 
-The distribution of the note are extremely similar.
+The distribution of the notations given by the subjects are extremely similar.
 The $p-value$ is calculated at $0.18 > \alpha$ which means that the two distribution cannot be distinguished.
 So the simulated scenes are perceived in a similar way than the recorded scenes.
-The mean for the simulated scene are even superior to the real ($m_{simul.} = 5.1 (\pm 1.6$), $ m_{real} = 4.9 (\pm 1.6$)).
+Although not significant, one can note that the mean for the simulated scene is even superior to the real ($m_{simul.} = 5.1 (\pm 1.6$), $ m_{real} = 4.9 (\pm 1.6$)).
 \\
 
-Secondly, an ANOVA is performed to determine if the experience in the listening of urban sound mixtures is an influential factor to distinguish real and replicated scenes.
+A second ANOVA is performed to determine if the experience in the listening of urban sound mixtures is an influential factor to distinguish real and replicated scenes.
 It is then a 2 dimensions ANOVA which is used.
-The test allow to determine  the influence of each dimension separately and their interaction.
-In this case, it means that the comportment of a dimension is depending on those of the other dimension.
+The test allows us to determine  the influence of each dimension separately and their interaction.
+In this case, it means that the behavior of a dimension is depending on those of the other dimension.
 The $p-value$ associated is submitted on the two-hypotheses test (table~\ref{tab:p_value_type_exp}).\\
@@
 
-A last ANOVA is performed with 2 dimensions to see if the scenes according to the \textit{type} and their \textit{sound environment} are evaluated differently (figure~\ref{fig:boxplot_type_ambience} and table \ref{tab:p_value_type_ambience}).
+A last ANOVA is performed with 2 dimensions to check if the scenes according to the \textit{type} and their \textit{sound environment} are evaluated differently (figure~\ref{fig:boxplot_type_ambience} and table \ref{tab:p_value_type_ambience}).
 
@@
 
-For the \textit{type} dimension, the effect is still not significance ($p-value > \alpha$) contrary to the \textit{sound environment} dimension.
-Thus, the realism of the listened scenes seems most influenced by the kind of sound environment than the type of sound scenes.
-As the mean of the \textit{noisy} and \textit{very noisy} atmospheres are higher, it seems that the presence of cars are a preponderant element to evaluate the realism of an audio mixture.
-To the opposite, for the \textit{quiet} and \textit{park} atmosphere, the evaluation is more disperse.
-From the comments let by the panelists, in the real scenes, the presence of too loud birds and some street noises with unknown origins have degraded the evaluation.
+For the \textit{type} dimension, the effect is still not significant ($p-value > \alpha$) contrary to the \textit{sound environment} dimension.
+Thus, the realism of the listened scenes seems to be more influenced by the kind of sound environment than the type of sound scenes.
+As the scores of the \textit{noisy} and \textit{very noisy} atmospheres are higher, it seems that the presence of cars are a preponderant element to evaluate the realism of an audio mixture.
+On contrary, for the \textit{quiet} and \textit{park} atmosphere, the evaluation is more complext.
+From the comments given by the subjects, in the real scenes, the presence of too loud birds and some street noises with unknown origins have degraded the evaluation.
 Whereas in the simulated scenes, it is mainly the sound class \textit{foot step} and a sound background composed of birds that made the realism of the scenes decrease.
@@
 \section*{Conclusion}
-Realistic urban sound mixtures have been composed from the study of urban recordings.
+
+In order to study the acoustic properties of the urban environnment, the research community designs tools that have to be evaluated to demonstrate their merit.
+We believe to the availability of simulated scenes with precise annotation in terms of centent and level that are realistic enough will foster research in this direction.
+
+To do so, realistic urban sound mixtures have been composed from the study of urban recordings.
 A work of annotation has allowed to extract, for different acoustic atmosphere, some useful information as the sound classes presents, the traffic flow rates, the sound level.
 Form these observation, a urban sound database have been created including car passages recordings.
-A perceptual test has been set up to prove the realism of the sound mixtures.
-This test revealed that the replicated scenes are mistaken with the urban recordings.
-Furthermore, the results shows that it is more the sound environment which impact the realism than the experience of the listener.
-In the \textit{park} and \textit{quiet street} environment, the evaluation are more disperses and the mean note lower than the \textit{noisy} and \textit{very noisy} atmosphere.
-Theses differences come from mainly from the sound level of some events which are too loud to be realistic enough but that can be improved.
+A perceptual test has been set up to quantify the level of realism of the simulated scenes.
+It showed that the level of realism of the simulated scenes is comparable to the one of the recorded scenes.
+
+Furthermore, the results shows that an important factor to consider is the sound environment impacts wrt.
+the experience of the listener.
+In the \textit{park} and \textit{quiet street} environment, the evaluation are more spread and the mean note lower than the \textit{noisy} and \textit{very noisy} atmosphere.
+Theses differences come from mainly from the sound level of some events which are found to be too loud to be realistic enough.% but that can be improved.
 From theses issues, it is possible to improve again the realism  of the replicated scenes by taking these comments account.

### 2017-06-19 articleBoston.tex
add boston-presentation

--- 2017-04-24 articleBoston.tex
@@
-\documentclass[a4,11pt,twocolumn]{article}
+\documentclass[11pt,letter]{article}
 
-\input{packageBoston}
+\usepackage{poma_style}
 
-\title{Design of a urban sound scenes corpuses with controlled acoustic properties}     %% \title est une macro, entre { } figure son premier argument
-\author{
-  GLOAGUEN, Jean-R\'emy\      \texttt{first1.last1@xxxxx.com}
-  \and
-  CAN, Arnaud\      \texttt{first2.last2@xxxxx.com}
-  \and
-  LAGRANGE, Mathieu\      \texttt{first1.last1@xxxxx.com}
-  \and
-  PETIOT, Jean-Fran\c cois\      \texttt{first1.last1@xxxxx.com}
-}
+\input{packageStyle}
+
+\newcommand{\eps}{\varepsilon}
 
@@
 
-\maketitle
+%\section*{Abstract}
+%
+%Monitoring the acoustic urban environment is beneficial for improving the life of citizens.
+To do so, the research community in acoustics can benefit from corpora where the acoustic contribution of sources of disturbance such as road traffic is precisely known.
+%
+%In this paper, a set of urban sound mixtures are simulated using an open source tool \textit{simScene}.
+Their characteristics (type of sources, density of events, etc.) are inferred from the listening of reference audio scenes recorded in Paris, France.
+The main advantage of the simulated versions is that several quantities (such as the acoustic level of each source, a useful feature automated analysis tools) is precisely known.
+%
+%In order to quantify the level of realism achieved by the use of the simulation tool a perceptive test is conducted with 50 subjects.
+The results show that no significant differences can be made between the simulated and the recorded scenes.
 
-\section*{Abstract}
+\section{Introduction}
 
-Monitoring the acoustic urban environment is beneficial for improving the life of citizens.
-To do so, the research community in acoustics can befenit from corpora where the acoustic contribution of sources of disturbance such as road traffic is precisely known.
+The urban sound environment is studied both from the perceptual \cite{jin_yong_soundwalk_2013} \cite{botteldooren_understanding_2011} and from the physical point of view \cite{can_describing_2015} \cite{raimbault_ambient_2003}.
+These studies can be based on real sound environments either by acoustic listening through a \textit{soundwalk} \cite{adams_soundwalking_2008} or with recordings made in city \cite{botteldooren_temporal_2006}, or in a laboratory by using sound mixtures resulting from a simulation process \cite{lafay_new_2014}.
+In the first case, since it has an undeniable ecological validity, it doesn't make it possible to have a repeatable experience nor to propose a controlled framework where the presence and the level of the different sound sources can be adjusted.
+Then the use of a simulation tool to create specific urban sound environments is a useful way to control the degree of contribution of each class of acoustical sources.
+It allows us to create urban sound environments by adjusting some parameters \cite{bruce_development_2009} or to evaluate performances of some classification or detection tools \cite{giannoulis_detection_2013}.
 
-In this paper, a set of urban sound mixtures are simulated using an open source tool \textit{simScene}.
-Their caracteristics (type of sources, density of events, etc.) are infered from the listening of reference audio scenes recorded in Paris, France.
-The main advantage of the simulated versions is that several quantities, such as the acoustic level of each source is precisely known, a useful feature for evaluated automated analysis tools.
-
-In order to quantifiy the level of realism achieved by the use of the simulation tool a perceptive test is conducted with 50 subjects.
-The result show that no significative differences can be made between the simulated and the recorded scenes.
-
-\section*{Introduction}
-
-The urban sound environment is studied both from the perception \cite{jin_yong_soundwalk_2013} \cite{botteldooren_understanding_2011} and from the physical point of view \cite{can_describing_2015} \cite{raimbault_ambient_2003}.
-These studies can be based on real sound environments either by acoustic listening through a \textit{soundwalk} \cite{adams_soundwalking_2008} or with recording made in city \cite{botteldooren_temporal_2006} or in a laboratory by using sound mixtures resulting from a simulation process \cite{lafay_new_2014}.
-In the first case, if the urban sound environment have an undeniable ecology validity, it doesn't make it possible to have a repeatable experience nor to propose a controlled framework where the presence and the level of the different sound sources can be adjusted.
-Then the use of a simulation tool to create specific urban sound environments is a useful way to control the degree of contribution of each class of acoustical sources.
-It allows us to create urban sound environments by adjusting some parameters \cite{bruce_development_2009} or to evaluate the performance of some classification or detection tools \cite{giannoulis_detection_2013}.
-
-A critical question is then to obtain scenes that are realistic enough, thus to be considered as similar to some reference recordings.
+A critical question is then to obtain scenes that are realistic enough to be considered simulated scenes as similar to some reference recordings.
 To do so, various methods have been proposed.
-The first approach would be to simulate completely a neighborhood taking into account the architecture of the buildings, the sound dynamic of the different sources present (car, voice, bird, bell \dots) and their propagation to a receptor, as proposed by \cite{cstb_simulation_2015}.
+The first approach would be to simulate completely a neighborhood taking into account the architecture of the buildings, the sound dynamics of the different sources present (car, voice, bird, bell \dots) and their propagation to a receptor, as proposed by \cite{cstb_simulation_2015}.
 If the first renderings are interesting, this method remains complex to implement.
-A more straitghtforward approach composes sound mixtures from isolated sound mixed together.
+A more straightforward approach composes sound mixtures from isolated sounds mixed together.
 The idea is to consider the urban sound environment as the sum of acoustic events, i.e.
 events with a sufficient sound level to be discernible, superposed to a sound background, i.e a sound constant on the sound mixture whose properties vary slowly during the scene.
-The difficulty here is to have a representative database of isolated sounds with a sufficient quality not to compomise the resulting sound quality and to sequence the acoustic events wisely.
+The difficulty here is to have a representative database of isolated sounds with a sufficient quality to not compromise the resulting sound quality and to sequence the acoustic events wisely.
 
-The method proposed in \cite{misra_musical_2007} enables to resolve the first issue by extracting the acoustical event directly in real recordings and by manipulating them to re-use them in new mixtures.
-If the tool has a sufficient number of parameters to control the sound event, their method is limited by the extraction phase: the overlapping between the acoustics events deteriorates the sounds and then the final rendering.
-In more, the manipulation (duration, frequency range) can create artifacts that can be a limitation for a realistic use.
+The method proposed in \cite{misra_musical_2007} enables to resolve the first issue by extracting the acoustical events directly in real recordings and by manipulating them to re-use them in new mixtures.
+Since the tool has a sufficient number of parameters to control the sound event, their method is limited by the extraction phase: the overlapping between the acoustic events deteriorates the sounds and then the final rendering.
+Furthermore, the manipulations (duration, frequency range) can create artifacts that can deacrease the realism.
 In a similar way, the tool proposed by Davis and Bruce \cite{bruce_development_2009} enables to compose sound mixtures with an isolated sound database composed of recordings made specially for their application.
 The creation of the urban sound mixtures is made possible with the perceptual evaluation of a panel of listeners of the different sound environments.
-This method enables to determine which sounds are the most representative or the most linked to an urban sound environment from a perspective point of view.
+This method enables to determine which sounds are the most representative or the most linked to an urban sound environment from a perceptual point of view.
 
-However, as our objective here is to get simulated scenes that are as realistic as possible, this method can not be considered because if this method proposes a perspective validity it can not provide an ecological one.
+However, as our objective here is to get simulated scenes that are as realistic as possible, this method can not be considered because if this method proposes a perceptual validity, it can not provide an ecological one.
 Thus, this study proposes to create urban sound mixtures from the listening of urban audio recordings and the estimation of some high level sequencing parameters extracted from the recordings to tune an audio simulation tool.
-The realism of the simulated scenes is then evaluated using a perspective test.
+The realism of the simulated scenes is then evaluated using a perceptual test.
 
-The remaining of the paper is organized as follows, Section \ref{sec:study} focuses on the study of recordings made in Paris.
-Section \ref{sec:simulation} deals with the simulation tool and Section \ref{sec:test} summarizes the result obtained in the perspective test.
+The remaining of the paper is organized as follows: Section \ref{sec:study} focuses on the study of recordings made in Paris, Section \ref{sec:simulation} deals with the simulation tool and Section \ref{sec:test} summarizes the results obtained in the perspective test.
 
-\section{Study of realistic scenes} \label{sec:study}
-Real urban noise recordings are listened in details to determine the composition of typical urban sound environment in terms of sound sources, presence and level.
-The recordings considered in this study have been made on the \nth{13} district of Paris (France) as part of the GRAFIC project \cite{aumond_sound_2016} during a soundwalk which was designed to cover different types of urban sound environments.
-The walk was 2.1 km long, consisting of 19 stop points of 3 to 5 minutes of listening each (see figure \ref{fig:soundwalk}).
-It was roamed on two days (03/23/2015 and 03/30/2015) twice a day (on the morning and on the afternoon) and in one direction (from West to East, WE) and then in the other (from East to West, EW).
-The acquisition system used was a ASA SENS and was carried on the backpack of the operator which was walking with the participants.
-In the end, 74 audio pulse code modulation recordings sampled at 44,1 kHz are available and serve as a base for the study.
-More details can be found on  \cite{aumond_modelling_2017}.
-\\
+\section{Study of real scenes} \label{sec:study}
+Real urban noise recordings are listened in details to determine the composition of typical urban sound environments in terms of sound sources, presence and level.
+The recordings considered in this study have been made on the \nth{13} district of Paris (France) as part of the GRAFIC project \cite{aumond_sound_2016} during a soundwalk which was designed to cover different types of urban sound environments.\\
 
@@
 
-Each audio is fully annotated in terms of sound classes that are be present in a urban area, their recurrence and is labeled using the following types of sound environments (\textit{park}, \textit{quiet street} , \textit{noisy street}, \textit{very noisy street}) as proposed by \cite{rychtarikova_soundscape_2013} or \cite{can_describing_2015} (table~\ref{tab:classificationScene}).
+The walk was 2.1 km long, consisting of 19 stop points with a recording duration between 3 to 5 minutes (see figure \ref{fig:soundwalk}).
+It was roamed on two days (03/23/2015 and 03/30/2015), twice a day (on the morning and on the afternoon) and in one direction (from West to East, WE) and then in the other (from East to West, EW).
+The acquisition system used was a ASAsense and was carried on the backpack of the operator which was walking with the participants.
+In the end, 76 (4 $\times$ 19) audio files (sampled at 44,1 kHz) were available and serve as a basis for the study.
+More details can be found on \cite{aumond_modelling_2017}.
+Each file is fully annotated in terms of sound classes that are present in the urban area and their recurrence.
+It is labeled using the following types of sound environments (\textit{park}, \textit{quiet street}, \textit{noisy street}, \textit{very noisy street}) as proposed by \cite{rychtarikova_soundscape_2013} or \cite{can_describing_2015} (Table~\ref{tab:classificationScene}).
 
@@
 \multicolumn{1}{|l|}{\textbf{day}} & \textbf{journey}   & 1                        & 2                        & 3                        & 4                        & 5                        & 6                        & 7                        & 8                        & 9                        & 10                       & 11                       & 12                       & 13                       & 14                       & 15                       & 16                       & 17                       & 18                       & 19                       \\ \hline
-\multirow{2}{*}{\textbf{1}} & \textbf{EW} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} \\
-\cline{2-20}
- & \textbf{WE} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} &                          & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} &  \\
+\textbf{1} & \textbf{EW} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} \\
 \hline
-\multirow{2}{*}{\textbf{2}} & \textbf{EW} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} \\
-\cline{2-20}
- & \textbf{WE} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{5AB25A} \\ \hline
+\textbf{1} & \textbf{WE} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} &                          & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} &  \\
+ \hline
+\textbf{2} & \textbf{EW} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} \\
+\hline
+\textbf{2} & \textbf{WE} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{5AB25A} \\ \hline
 \end{tabular}
-
+%
 \vspace{0.5cm}
@@
 
-Most of the scene of the recordings belong to \textit{quiet street} and \textit{noisy street}.
-2 scenes are not available : a noisy cleaning truck is present throughout the \textit{1-WE-3} heavily polluting the sound scene and the \textit{1-WE-19} record is too short to be considered in this study.
-From the annotation, each sound ambiance is characterized with the sound classes presents in it.
-The most common sound classes are displayed in Table \ref{tab:obsScene}.\\
+Most of the scenes of the recordings belong to \textit{quiet street} and \textit{noisy street}.
+2 scenes are removed form the database: a noisy cleaning truck is present inside the \textit{1-WE-3} heavily polluting the sound scene and the \textit{1-WE-19} record is too short to be considered in this study.
+From the annotations, each sound environment is characterized by the sound classes present in the recordings.
+The most recurrent sound classes are presented in Table \ref{tab:obsScene}.\\
 
@@
 \end{tabular}
-\caption{Sound level and classes presents depending on the sound environment}
+\caption{Sound level and description of the most recurrent sound classes (backgrounds and events) present in the sound environments}
 \label{tab:obsScene}
@@
 
-%\begin{table*}[t]
-%\centering
-%
-%\begin{tabular}{p{3cm}p{2cm}p{10.5cm}}
-%\hline
-%\multicolumn{1}{c}{\textbf{Ambience}} &\textbf{\begin{tabular}[c]{@{}c@{}}level\\ (dB SPL)\end{tabular}} & \multicolumn{1}{c}{\textbf{Description}} \\ \hline
-%park & \centering 69,0 & predominance of children's and adults' voices and bird whistles as sound events and backgrounds, the road traffic is low almost non-existent ($\approx$ 1 car/min) \\
-%\hline
-%quiet street & \centering 70,2 & The background noise is caused by a low road traffic with few car passages ($\approx$ 2 car/min), the birds are present and audible, the voice is composed of children's and adult's voices.
-The effect of reverberation on the street, due to low background noise, is present.\\
-%\hline
-%animated street & \centering 73,5 & traffic noise  is louder and the unique background, the car passages density is higher ($\approx$ 10 car/min), the human activity is more present and diversify (voice, step, suitcase, construction site noise \dots), the birds disappear.
-\\
-%\hline
-%very noisy street & \centering 76,0 & the background noise very high composed of quasi-constant traffic flow ($\sup$ 40 car/min), the human activity are less presents in a favor of voice hubbub.\\
-%\hline
-%
-%\end{tabular}
-%\caption{Qualitative observations on the urban sound environment}
-%\label{tab:obsScene}
-%\end{table*}
+As the calibration audio file was not available, the sound level cannot be considered as absolute but the relative difference between each sound environment can be considered.
+The road traffic, the voice and the bird components are the most discriminant classes of the sound environment: with exclusively road traffic, the \textit{noisy} and \textit{very noisy} environments sound environments are different to the \textit{park} and \textit{quiet street} where the voice and the bird's whistles are predominant.
+Furthermore, multiple sound classes can be heard whatever the urban sound environments : dog barks, church bell ringing, car horn, sirens \dots with a density that can be very low (< 0.1/min).
+Finally, a lot of brief sounds with unknown origin can be listened in almost all the files.
+Consequently, all these indefinite sounds are annotated in one unique class sound, \textit{street noise}.
+From this information (density, level, sound classes present in each sound environment), it is now possible to compose urban sound mixtures.
 
-The road traffic, the voice and the bird components are the most linked to the sound atmosphere: with exclusively road traffic, the sound environment is more noisy than in the \textit{park} and \textit{quiet street} where the voice and the bird's whistles are predominant.
-Furthermore, multiple sounds classes can be heard whatever the urban sound environments : dog barks, church bell ringing, car horn, sirens \dots with a density which can be very low (< 0.1/min).
-Finally, a lot of brief mechanics sounds with unknown origin can be listened in almost all the audio.
-Consequently, all these indefinite sounds are annotated in one unique class sound, \textit{street noise}.
-From theses information (density, level, sound classes present in each ambience), it is now possible to compose urban sound mixtures.
-
-\section{Simulation of realistic simulated urban scenes} \label{sec:simulation}
+\section{Simulation of realistic auditory urban scenes} \label{sec:simulation}
 \subsection{Presentation of the web-simulator \textit{simScene}}
-\textit{simScene} \cite{rossignol_simscene:_2015} is a simulator that creates sound mixtures in a wav format by superposing audio which comes from an isolated sound database composed of two categories: the brief sounds (from 1 to 20 seconds)
-that are considered as salient belong to the \textit{event} category and all the sound that are of long duration and whose acoustic properties do not vary with respect to time belong to the \textit{background} category.
-In each category, the sounds are then grouped in sound classes (bird, car, foot steps \dots).
-The software enables the user to control for each sound classes some parameters as the number of events of each class that appears in the mixture, the elapsed time between each sample of a same class or the existence of a fade in and a fade out \dots.
-Each parameters are completed with a standard deviation optionally bringing some random behavior between the scenes.
+\textit{simScene} \cite{rossignol_simscene:_2015} is a simulator that creates sound mixtures in a .wav format by superposing audio samples that come from an isolated sound database composed of two categories: the brief sounds (from 1 to 20 seconds) that are considered as salient belong to the \textit{event} category whereas all the sound that are of long duration and whose acoustic properties do not vary with respect to time belong to the \textit{background} category.
+Inside each category, the sound samples are then grouped in sound classes (bird, car, foot steps \dots).
+The software enables the user to control for each sound class some parameters as the number of events of each class that appears in the mixture, the elapsed time between each sample of a same class or the existence of a fade in and a fade out \dots.
+Each parameter is completed with a standard deviation that may brings some random behavior between the scenes.
 \\
@@
 \centering
-\includegraphics[width = \linewidth]{./pictures/spectrogramme_abstract_sceneSimpleKlaxonFixe_EN.pdf}
+\includegraphics[width = .7\linewidth]{./pictures/spectrogramme_abstract_sceneSimpleKlaxonFixe_EN.pdf}
 \caption{Example of a scene composed of three sound classes}
@@
 
-The sound mixtures can be created following two modes: the \textit{abstract} mode allows us to create mixtures from the specified parameters while the \textit{replicate} mode reproduces an existing scene by following as much as possible its annotation file.
-With the global sound mixture, an audio can also be created for each sound class present in the scene which allows to know their exact contribution in the scene.
-%In parallel, \textit{simScene} generates 3 image files representing the temporal evolution, the spectrogram and a piano Roll representation to visualize each sound classes distribution and also two text files summarizing the temporal location of each sample and all the parameter used.
-\\
+The sound mixtures can be created following two modes: the \textit{abstract} mode allows to create mixtures from the specified parameters while the \textit{replicate} mode reproduces an existing scene by following as much as possible its annotation file.
+With the audio file of the global sound mixture, each sound class can be resumed in a separate audio file, which allows to know their exact contribution in the scene.
 
 \subsection{Creation of the urban sound database}
-Based on the annotations of the real scenes, a sound database is built using the \textit{replicate} mode with sounds found online on the \textit{freesound} project and with the urbanSound8k database\cite{salamon_dataset_nodate}.
-This database consists in more than 8000 files with a 4 seconds or less duration found on the web site freesound.org too, and classified in 10 sound classes : ventilation, car horn, children playing, dog barking, bell, engine idling, gunshot, jackhammer, siren and street music.
-% ML Pas clair, est-ce utile ?
-All the audio have been sorted and those who presented the best signal noise ratio are selected in the sound classes identified in the real records have been selected.\\
+Based on the annotations of the 74 real scenes, a sound database is built using the \textit{replicate} mode of \textit{simScene} with sounds found online on the \textit{freesound} project and with the help of the urbanSound8k database \cite{salamon_dataset_nodate}.
+This database consists in more than 8000 files with a 4 seconds or less duration, found on the web site freesound.org too, and classified in 10 sound classes : ventilation, car horn, children playing, dog barking, bell, engine idling, gunshot, jackhammer, siren and street music.
+All the audio files have been sorted and those who presented the best signal to noise ratio have been selected.\\
 
-In addition, as road traffic is a prime audio source in an urban sound environment, it seemed interesting to have, in the database, a part composed of well-recorded car audio.
-As a result, passages of 4 different cars (Renault Scenic, Clio, Megane and Dacia Sandero) were recorded on the Ifsttar-Nantes runway at different speeds and gear ratios for steady, acceleration and braking phases (table~\ref{tab:resum_car_audio}).
+In addition, as road traffic is a prime audio source in an urban sound environment, it seemed interesting to have, in the database, a part composed of well-recorded car sounds.
+As a result, passages of 4 different cars (Renault Scenic, Clio, Megane and Dacia Sandero) were recorded on the Ifsttar-Nantes runway at different speeds and gear ratios for steady, acceleration, braking (Table~\ref{tab:resum_car_audio}) and stopped phases.
+Among the 108 recordings planned, those for the acceleration and braking phase for the Renault Senic could not be done whereas some recordings have been made twice for the other vehicles.
 In all, 103 car passages have been recorded.
@@
 \multicolumn{2}{|c||}{\textbf{braking}} & \multicolumn{2}{|c|}{\textbf{acceleration}} \\ \hline
-\multicolumn{1}{|c|}{\begin{tabular}[c]{@{}c@{}}\textbf{speed}\\  \textbf{(km/h)}\end{tabular}} & \multicolumn{1}{c||}{\textbf{trans.}} & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}\textbf{speed}\\  \textbf{(km/h)}\end{tabular}} & \multicolumn{1}{c|}{\textbf{trans.}} \\ \hline
-50 → 0 & 3 → 2 & 0 → 30 & 1 → 2 \\ \hline
-40 → 0 & 2 → 2 & 0 → 40 & 1 → 2 \\ \hline
-50 → 30 & 3 → 2 & 20 → 40 & 1 → 3 \\ \hline
-60 → 40 & 4 → 3 & 30 → 50 & 2 → 3 \\ \hline
-70 → 50 & 4 → 3 & 40 → 60 & 3 → 4 \\ \hline
-80 → 50 & \begin{tabular}[c]{@{}l@{}}4 or 5\\ → 3\end{tabular} & 50 → 70 & \begin{tabular}[c]{@{}l@{}}3 → \\ 4 or 5\end{tabular} \\ \hline
+\multicolumn{1}{|c|}{\begin{tabular}[c]{@{}c@{}}\textbf{speed}\\  \textbf{(km/h)}\end{tabular}} & \multicolumn{1}{c||}{\textbf{gear r.}} & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}\textbf{speed}\\  \textbf{(km/h)}\end{tabular}} & \multicolumn{1}{c|}{\textbf{gear r.}} \\ \hline
+50 $\rightarrow$ 0 & 3 $\rightarrow$ 2 & 0 $\rightarrow$ 30 & 1 $\rightarrow$ 2 \\ \hline
+40 $\rightarrow$ 0 & 2 $\rightarrow$ 2 & 0 $\rightarrow$ 40 & 1 $\rightarrow$ 2 \\ \hline
+50 $\rightarrow$ 30 & 3 $\rightarrow$ 2 & 20 $\rightarrow$ 40 & 1 $\rightarrow$ 3 \\ \hline
+60 $\rightarrow$ 40 & 4 $\rightarrow$ 3 & 30 $\rightarrow$ 50 & 2 $\rightarrow$ 3 \\ \hline
+70 $\rightarrow$ 50 & 4 $\rightarrow$ 3 & 40 $\rightarrow$ 60 & 3 $\rightarrow$ 4 \\ \hline
+80 $\rightarrow$ 50 & \begin{tabular}[c]{@{}l@{}}4 or 5\\ $\rightarrow$ 3\end{tabular} & 50 $\rightarrow$ 70 & \begin{tabular}[c]{@{}l@{}}3 $\rightarrow$ \\ 4 or 5\end{tabular} \\ \hline
 \end{tabular}
     \end{minipage}%
-    \caption{Measurement set on runway with vehicle passages with stabilized speed (left) and in acceleration and braking phase(right)}
+    \caption{Description of the recordings set on runway with vehicle passages with stabilized speed (left) and in acceleration and braking phase(right) with different gear ratios}
     \label{tab:resum_car_audio}
@@
 
-A median filter \cite{fitzgerald_harmonic/percussive_2010} has been applied to filter out bird's whistles from the recordings in the frequency range $\left[ 2500-6500 \right]$ Hz.
-It   consists in taking a window around a center point and to attribute to it the median value of this window.
-This method allow to attenuate the brief variations both in the harmonic and the temporal plan.
-The size of the filter window is chosen with a 5 points width of and a 9 points height so that the effect of the filter is sufficient without deteriorating the quality of the audio (see figure \ref{fig:filtre_car}).
+To obtain the cleanest possible sound samples, a median filter \cite{fitzgerald_harmonic/percussive_2010} has to be applied to filter out bird's whistles present in the recordings.
+The filter consists in taking a window around a center point and to attribute to it the median value of this window.
+This method allows to attenuate the brief variations both in the spectral and the temporal plan.
+The size of the filter window is chosen with a 5 points width and a 9 points height in the frequency range $\left[ 2500-6500 \right]$ Hz, so that the effect of the filter is sufficient without deteriorating the quality of the audio (see Figure \ref{fig:filtre_car}).
 
@@
 \centering
-\includegraphics[width=\linewidth]{./pictures/filtrageMedian_VL1_R3_40_EN.pdf}
-\caption{Spectrogram of a recorded passage of the Renault Megane car at 40 km/h with the \nth{3} gear ratio ($N_{w} = 2^{12}$ with 50 $\%$ overlapping, $N_{fft} = 2^{12}$, $sr = 44,1$ kHz).
-On the left, the the original recording, on the right, the filtered audio with the filtered part framed.}
+\includegraphics[width=.7\linewidth]{./pictures/filtrageMedian_VL1_R3_40_EN.pdf}
+\caption{Spectrogram in the frequency range $\left[ 2500-6500 \right]$ Hz of a recorded passage of the Renault Megane car at 40 km/h with the \nth{3} gear ratio ($N_{w} = 2^{12}$ with 50 $\%$ overlapping, $N_{fft} = 2^{12}$, $sr = 44.1$ kHz).
+On the left, the original recording, on the right, the filtered audio with the filtered part.}
 \label{fig:filtre_car}
@@
 With this built-up database, the 74 real sound scenes are replicated.
-To test whether the replicated scenes are sufficiently realistic, several of them are evaluated by a panel of listeners.
+In order to obtain a corpus of urban sound scenes, resulting from a simulated process, sufficiently realistic, a perceptual test is set up to evaluate the realism on several of them.
 
 \section{Perceptual test} \label{sec:test}
-\subsection{Creation of the test}
-A perspective test is created where a panel of listeners evaluates the realism of a mix of real and replicated scenes on a 7-point scale (1 is \textit{not realistic at all}, 7 is \textit{very realistic}).
-Since it is not possible for each panelist to evaluate all the real and replicated scenes, in order to limit the duration of the test and to keep the judge's concentration constant, each listens a panel of 20 sound scenes composed of 10 real sounds scenes and 10 replicated.
-Furthermore, for the auditory comfort of the panelist, all the scenes are normalized to the same sound level, chosen at 65 dB.\\
+\subsection{Design of the test}
+A perceptual test is conducted with a panel of listeners that are asked to evaluate the level of realism on a 7-point scale (1 is \textit{not realistic at all}, 7 is \textit{very realistic}) of an auditory scenes, recorded and simulated.
+The total number of sound scenes to be tested is set at 40.
+It consists of 20 30-seconds audio files, including 5 scenes that belong to the sound environment \textit{Park}, 6 from \textit{Quiet street}, 4 from \textit{Noisy street} and 5 from \textit{Very noisy street} chosen randomly among the 74 recorded scenes, and the same 30 seconds from the replicated scenes.
+In order to limit the duration of the test and to preserve the concentration of the subjects, each subject listens a subset of 20 sound scenes (10 real sounds scenes and 10 replicated).
+Furthermore, to prevent the listeners from changing the sound level of their speakers, all the scenes are normalized to the same sound level, chosen at 65 dB.\\
 
-First, the listening plan is elaborated following a Balanced Incomplete Block Design \cite{dagnelie_principes_2003} where $J$ participants realized $K$ evaluations from $B$ sound sources, each of them are tested $R$ times and each couple of sound sources are presented $\lambda$ times.
-3 rules govern these parameters :
+The experimental design is elaborated following a Balanced Incomplete Block Design (BIBD) \cite{dagnelie_principes_2003} where $J$ participants make $K$ evaluations from $B$ sound scenes, each of them are tested $R$ times and each couple of sound sources are presented $\lambda$ times in one block.
+In order to have BIBD, a 3 rules govern these parameters :
+
 \begin{subequations}\label{eq:BIE_cond}
 \begin{align}
-J &\geq B, \label{eq:BIE_cond1}\\
+B &\geq K, \label{eq:BIE_cond1}\\
 JK &= BR, \label{eq:BIE_cond2}\\
-\lambda &= R\frac{K-1}{J-1}. \label{eq:BIE_cond3}
+\lambda &= R\frac{K-1}{B-1}. \label{eq:BIE_cond3}
 \end{align}
@@
 with $\left[J, B, K, R, \lambda\right] \in \mathbb{N}$.
-A reasonable number of listeners is chosen at $J = 50$ and the total numberof sound sources $B$ to be tested is set at 40.
-It consists of 20 30-seconds audios from the recorded scene and the same 30 seconds from the replicated scenes.
-Among the 20 scenes, there are 5 scenes which belong to the ambience \textit{Park}, 6 from \textit{Quiet street}, 4 from \textit{noisy street} and 5 from \textit{very noisy street} chosen randomly among the 74 scenes.
-As (\ref{eq:BIE_cond3}) is not satisfied, the listening plan is finally elaborated following an optimal plan \cite{pages_blocs_2007}.
-It consists, with the parameters $J$, $K$ and $R$ fixed, to determine the optimal plan that would be the most balanced (figure \ref{fig:repartition}).\\
+A reasonable number of listeners is chosen at $J = 50$.
+With theses parameters fixed, $J=50$, $K=20$, $B=40$ and $R=25$, the BIBD is not fully balanced (condition (\ref{eq:BIE_cond3}) $\notin \mathbb{N}$).
+It is finally with a partially BIBD that the listening plan is elaborated \cite{pages_blocs_2007}.
+It consists, with the parameters $J$, $K$ and $R$ fixed, in determining the optimal plan that would be the most balanced.
+The package \textit{sensoMineR} on the \textit{R} software provides the plan where the listening order per participant is defined inside \cite{le_sensominer:_2008}.
+As the distribution of the simulated and real samples for each participant is almost identical, the obtained listening plan can be considered as balanced (Figure \ref{fig:repartition}).\\
 
@@
 \centering
-\includegraphics[width = \linewidth]{./pictures/repartition-real-simulated.pdf}
+\includegraphics[width = .7\linewidth]{./pictures/repartition-real-simulated.pdf}
 \caption{Distribution between the real and simulated scenes for each judge. The cumulative sum is equal to the number of tested elements $K$.}
@@
 
-A web-page\footnote{http://soundthings.org/research/xpRealism} was uploaded on the 8 february 2007 and the number of 50 compiled test has been reached 12 days later.
-During the test, the participant had the possibility to listen each scene as many time as wanted before evaluating it without being able to change his judgment afterwards.
-He could also leave a comment on each audio to explain his note.
-Finally, his age, gender and experience on listening to urban sound mixtures were asked at the end of the test.
-In the end, the panel of 50 listeners was made of 31 males and 18 females (one not documented) with an average age of 36 ($\pm$ 12) years old.
-$62\%$ of the participants declared they had no experience in the listening of urban sound mixtures.\\
+The test was administered online\footnote{\url{http://soundthings.org/research/xpRealism}} on the 8 February 2017 and the number of 50 participants has been reached 12 days later.
+During the test, the participant had the possibility to listen to each scene as many times as wanted before evaluating it, without being able to change his/her judgment afterwards.
+The participant could also leave a comment on each audio to explain the rating.
+Finally, the age, gender and experience on listening to urban sound mixtures were asked at the end of the test.
+The panel of 50 listeners was made of 31 males and 18 females (one not documented) with an average age of 36 ($\pm$ 12) years old.
+$62\%$ of the participants declared having no experience in the listening of urban sound mixtures.\\
 
@@
 
-An analyze of variance (ANOVA) is performed to see if both type of scenes can be discernible.
-2 hypothesis are then possible : $H_0$, the real and replicates scenes can not be discernible, and $H_1$, both scenes are perceived with differences.
-For that, an threshold significance $\alpha=0,05$ is compared to the \textit{p-value} calculated by the ANOVA.
-This value represents probability to have an extreme value or more than the actual data observed considering the $H_0$ hypothesis.
-Two situations can happen :
-\begin{itemize}
-\item $\alpha > p-value$, the $H_1$ hypothesis is validated and $H_0$ is rejected,
-\item $\alpha < p-value$, the $H_0$ is not rejected.
-\end{itemize}
-
-The data are represented by a box-and-whiskers plot according their type.
+First, the note distributions are represented by a box-and-whiskers plot according to whether they belongs to the type 'real' or 'simulated'. 
 \begin{figure}[hbtp]
 \centering
-\includegraphics[width=\linewidth]{./pictures/testPerceptif_boxplotType_EN.pdf}
-\caption{Box-and-whiskers plot according the type of scene}
+\includegraphics[width=.6\linewidth]{./pictures/testPerceptif_boxplotType_EN.pdf}
+\caption{Box-and-whiskers plot of the rating of realism according to the type of scene}
 \label{fig:boxplot_real_simul}
@@
 
-The distribution of the note are extremely similar.
-The $p-value$ is calculated at $0.18 > \alpha$ which means that the two distribution cannot be distinguished.
-So the simulated scenes are perceived in a similar way than the recorded scenes.
-The mean for the simulated scene are even superior to the real ($m_{simul.} = 5.1 (\pm 1.6$), $ m_{real} = 4.9 (\pm 1.6$)).
-\\
+The distribution allow to considered that the notation given the subjects are extremely similar.
+One can note that the mean for the simulated scene is even superior to the real one ($m_{simul.} = 5.1 (\pm 1.6$), $ m_{real} = 4.9 (\pm 1.6$)).
 
-Secondly, an ANOVA is performed to determine if the experience in the listening of urban sound mixtures is an influential factor to distinguish real and replicated scenes.
-It is then a 2 dimensions ANOVA which is used.
-The test allow to determine  the influence of each dimension separately and their interaction.
-In this case, it means that the comportment of a dimension is depending on those of the other dimension.
-The $p-value$ associated is submitted on the two-hypotheses test (table~\ref{tab:p_value_type_exp}).\\
+A paired samples \textit{t-test} is then performed to validate the similarity between the two types.
+It consists in validate (or not) an $H_0$ hypothesis which considers the similarity between the distribution of the average scores for the recorded and the simulated scenes of each judge with a Student test.
+This test establishes a \textit{p-value} that is compared to a threshold value $\alpha$ of $5\%$.
+The $H_0$ hypothesis is then considered if $p-value > \alpha$.
+The results of the \textit{t-test} (degrees of freedom (DOF), the absolute value \textit{t} and \textit{p-value}) are sum up in the table \ref{tab:ttest_type_of_scene}.
+
+\begin{table}[H]
+\centering
+\begin{tabular}{cccc}
+        & DOF & \textit{$\mid$ t $\mid$}    & \textit{p-value} \\ 
+\hline
+type & 49 & 1.37 &  0.17   \\
+\hline
+\end{tabular}
+\caption{T-test performed on the distribution of the average scores for the recorded and simulated scenes}
+\label{tab:ttest_type_of_scene}
+\end{table}
+
+The $p-value$ is calculated at $0.17 > \alpha$, which confirms that all the real and simulated scenes are perceived in a similar way by the panel.\\
+
+From this global result, a first analyze of variance (ANOVA) is performed to determine if the experience in the listening of urban sound mixtures is an influential factor to distinguish real and replicated scenes.
+In a similar way than the \textit{t-test}, but based on a Fischer statistical test, an ANOVA allows to determine if the distribution of different level of multiple factors are similar.
+It allows too to take into account multiple factors and their interactions.
+Therefore, a two-way ANOVA with interaction with the factors 'experience of the panelist in the listening of urban sound scene' (yes or no) and 'type of scene' (real or simulated) is carried out.
+The DOF, the Fisher statistics and the associated \textit{p-value} are given in Table~\ref{tab:ANOVA_type_exp}.\\
 
@@
 \begin{tabular}{cccc}
-\cline{2-4}
-          & type & experience & interaction \\
+        & DOF & \textit{F-statistic}    & \textit{p-value} \\ 
 \hline
-$p-value$ & 0.50 & 0.05 & 0.17 \\
+type & 1 & 0.59 & 0.44 \\
+\hline
+experience  & 1 & 1.97 & 0.16 \\ 
+\hline
+type/exp. & 1 & 2.14 & 0.14 \\
 \hline
 \end{tabular}
-\caption{$p-value$ for the ANOVA at two dimensions with the interaction effect}
-\label{tab:p_value_type_exp}
+\caption{Two-ways ANOVA with the factors 'type of sound' and 'listener experience'.}
+\label{tab:ANOVA_type_exp}
 \end{table}
 
-The dimensions and their interaction are not significantly influential since their $ p-value $ is equal to or greater than the $ \alpha$ significance threshold.
-This means that even listeners trained to hear urban sound environments do not dissociate the real scenes from the simulated ones.\\
+The effect of each factors and the interaction between them are not significant (\textit{p-value} $>$ 0.05).
+This means that even listeners trained to hear urban sound environments do not dissociate the real scenes from the simulated scenes better than the non-experienced listeners.\\
 
-A last ANOVA is performed with 2 dimensions to see if the scenes according to the \textit{type} and their \textit{sound environment} are evaluated differently (figure~\ref{fig:boxplot_type_ambience} and table \ref{tab:p_value_type_ambience}).
+The effect of the 'scene type' and the 'sound environment' factors and the interaction are then studied in a two-ways ANOVA (data distributions displayed on Figure \ref{fig:boxplot_type_ambience} and results in Table \ref{tab:p_value_type_ambience}).\\
 
@@
 \centering
-\includegraphics[width=\linewidth]{./pictures/testPerceptif_boxplotAmbianceCOLOR_EN.pdf}
-\caption{Box-and-whiskers plot according the type and the sound environment}
+\includegraphics[width=.7\linewidth]{./pictures/testPerceptif_boxplotAmbianceCOLOR_EN.pdf}
+\caption{Data distribution for the 'scene type' and 'sound environment' factors}
 \label{fig:boxplot_type_ambience}
@@
 \begin{tabular}{cccc}
-\cline{2-4}
-          & type   & sound env. & interaction \\
+        & DOF & \textit{F-statistic}    & \textit{p-value} \\ 
 \hline
-$p-value$ & $0.13$ & $0.80 \times 10^{-3}$     & $2.20 \times 10^{-3}$ \\
+type & 1 & 1.38 & 0.24   \\
+\hline
+sound env.   & 3 & 4.69 & 3.60 $\times 10^{-3}$  \\ 
+\hline
+judge.   & 49 & 5 & 0.18 $\times 10^{-9}$  \\ 
+\hline
+type/sound env. & 3 & 6.80 & 0.20 $\times 10^{-3}$\\
+\hline
+type/judge & 49 & 1.07 & 0.35\\
+\hline
+sound env./judge & 147 & 1.43 & 1.60 $\times 10^{-3}$\\
 \hline
 \end{tabular}
-\caption{$p-value$ for the anova at two dimensions with the interaction effect}
+\caption{Two-ways ANOVA for the 'type of sound' and 'sound environment' factors with the interaction effect}
 \label{tab:p_value_type_ambience}
@@
 
-For the \textit{type} dimension, the effect is still not significance ($p-value > \alpha$) contrary to the \textit{sound environment} dimension.
-Thus, the realism of the listened scenes seems most influenced by the kind of sound environment than the type of sound scenes.
-As the mean of the \textit{noisy} and \textit{very noisy} atmospheres are higher, it seems that the presence of cars are a preponderant element to evaluate the realism of an audio mixture.
-To the opposite, for the \textit{quiet} and \textit{park} atmosphere, the evaluation is more disperse.
-From the comments let by the panelists, in the real scenes, the presence of too loud birds and some street noises with unknown origins have degraded the evaluation.
-Whereas in the simulated scenes, it is mainly the sound class \textit{foot step} and a sound background composed of birds that made the realism of the scenes decrease.
-\\
+%Again, both the effects of each factors are not significant ($p-value > \alpha$) in contrast to its interaction ($p-value < \alpha$).  \\
 
-\section*{Conclusion}
-Realistic urban sound mixtures have been composed from the study of urban recordings.
-A work of annotation has allowed to extract, for different acoustic atmosphere, some useful information as the sound classes presents, the traffic flow rates, the sound level.
-Form these observation, a urban sound database have been created including car passages recordings.
-A perceptual test has been set up to prove the realism of the sound mixtures.
-This test revealed that the replicated scenes are mistaken with the urban recordings.
-Furthermore, the results shows that it is more the sound environment which impact the realism than the experience of the listener.
-In the \textit{park} and \textit{quiet street} environment, the evaluation are more disperses and the mean note lower than the \textit{noisy} and \textit{very noisy} atmosphere.
-Theses differences come from mainly from the sound level of some events which are too loud to be realistic enough but that can be improved.
-From theses issues, it is possible to improve again the realism  of the replicated scenes by taking these comments account.
+Again, the 'type' factor is not influential in the distribution of the notes ($p-value > \alpha$) whereas the 'judge' and the 'sound environment' factors are.
+But most of all, interactions effects occur for the sound environment with respectively the 'type' factor and the 'judge' factors.
+The interaction effect can be easily illustrated for the case of the type and sound environment factors (figure \ref{fig:interaction}).
+It reveals that the perception of realism of the simulated and real scenes is not the same depending on the sound environment.
 
-\section*{Acknowledgements}
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=.5\linewidth]{pictures/testPerceptif_interactionAmbianceCOLOR_EN.pdf}
+\caption{Interaction of the two factors 'type of sound' and 'sound environment'}
+\label{fig:interaction}
+\end{figure}
+
+For the \textit{quiet} and \textit{noisy} sound environments, the mean scores increase when the scenes are simulated whereas for the \textit{park} and \textit{very noisy} sound environment it decreases.
+As the effect of the \textit{type} factor is not the same according to the \textit{sound environment} levels, there is therefore an interaction effect.
+One notices that, on the simulated scenes, as the average scores of the \textit{noisy} and \textit{very noisy} atmospheres are higher for both, it seems that the presence of cars has been a preponderant element to increase the perceived realism of an audio mixture.
+On the contrary, for the \textit{quiet} and \textit{park} atmospheres, the evaluation is more complex.
+From the comments given by the subjects, it is mainly the sound class \textit{foot step} and a sound background composed of birds that have decreased the scene scores.
+In a similar way, in the real scenes, it is the presence of too loud birds and some street noises with unknown origins that have been remarked by the panelist and have degraded the evaluation.\\
+
+\section{Conclusion}
+
+In order to study the acoustic properties of the urban environment, the research community designs tools that have to be evaluated to demonstrate their merit.
+The availability of simulated scenes realistic enough with precise annotation in terms of content and level will foster research in this direction.
+
+To do so, realistic urban sound mixtures have been composed from the analysis of urban recordings.
+A work of annotation has allowed to extract, for different acoustic atmospheres, some useful information as the sound classes presence, the traffic flow rates, the sound level.
+From these observations, a urban sound database has been created including multiple sound classes both for the sound events and backgrounds.
+As the road traffic is one of the most noise source, to have a full control on this, multiple car passages recordings have been made on a runway.
+A perceptual test has been set up to quantify the level of realism of the simulated scenes and has demonstrated that they are comparable to the recorded scenes.
+Then, the different ANOVA performed have demonstrated that, according to the judge and, most of all, to the sound environment factors, they may still be significant differences in the distribution of notes.
+In the \textit{park} and \textit{quiet street} environment, for the simulated scenes, the evaluations are more spread and the mean notes lower than the \textit{noisy} and \textit{very noisy} atmospheres.
+Theses differences come mainly from the sound level of some events which are found to be too loud to be realistic enough.
+
+This problem solved, databases of sound sources and urban scenes can be shared with communities interested in methods of recognizing or detecting sound sources in order to help them to develop methods based on urban sound environments.
+
+\section{Acknowledgements}
 We would like to thank Pierre Aumond and Catherine Lavandier from the University of Cergy-Pontoise for transmitting us the data of the \textit{Grafic} project.

### 2017-06-19 boston-presentation.tex
add boston-presentation

\documentclass{beamer}

\input{packageLang}
%\input{../packageStyle}
\usepackage{adjustbox}
\usepackage{tabularx}
\usepackage{nth}
\usepackage[backend=bibtex]{biblatex}
\usepackage{caption}
\usepackage{media9}

%\bibliography{bibliography}

\setbeamertemplate{blocks}[rounded][shadow=true]
\usetheme{Boadilla}

\titlegraphic{
\begin{center}
\includegraphics[height=10mm]{./pictures/Ifsttar_logo_coul.jpg}
\hspace*{0.4cm}~
\includegraphics[width=10mm]{./pictures/LS2N.jpg}
\hspace*{0.5cm}~
\includegraphics[height=7mm]{./pictures/irstv_cle01beec-1.jpg}
\hspace*{0.4cm}~
\includegraphics[height=7mm]{./pictures/Region_Pays-de-la-Loire.png}
\end{center}
}

\title[Acoustics'17 Boston]{Creation of a corpus of realistic urban sound scenes with controlled acoustic properties}
\author[J. Gloaguen]{\emph{Jean-R\'emy Gloaguen}$^1$,  Arnaud Can$^1$,\\ Mathieu Lagrange$^2$,  Jean-Fran\c cois Petiot$^2$}
\institute[Ifsttar]{$^1$French Institute of Science and Technology for Transport, Development and Networks, \\ Environmental Acoustics Laboratory \\Allée des Ponts et Chaussées, Route de Bouaye, - CS4, 44344 Bouguenais Cedex, FR  \\ \vspace{.25cm} $^2$Nantes' Numeric Sciences Laboratory, \\1, rue de la Noë, BP 92101, 44321 Nantes Cedex 3, FR}
\date[06/29/2017]{Acoustics' 17 Boston\\06/29/2017}

\begin{document}

\begin{frame}

\titlepage

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\section{Introduction}
\frametitle{Introduction}

\begin{block}{Urban sound environments}

\begin{itemize}
	\item Subject of many researches both on perceptual and physical aspects
	\item Based on real sound environments with \textit{soundwalks} or recordings listened on laboratory\\
	$\Rightarrow$ ecological validity \\
	$\Rightarrow$ no controlled framework \\
	$\Rightarrow$ no repeatable experience \\

	\item Based on sound mixtures based on simulation process

		$\Rightarrow$ controlled framework  \\
		$\Rightarrow$ creation of specific urban sound environments \\
		$\Rightarrow$ no ecological validity \\
\end{itemize}

\end{block}
\vspace{0.5cm}
How can we simulate sound mixtures of urban sound environments that are realistic enough ?

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[label=pagebanale]
\frametitle{Introduction}

\begin{block}{Different approaches}
\begin{itemize}
	\item Simulate an entire neighborhood (buildings, sound emissions and propagation)
	\begin{itemize}
		\item too heavy to implement\\
		\item still too artificial \\
	\end{itemize}

	\item Design sound mixtures by superposing sound events with sound backgrounds
	\begin{itemize}
		\item build a representative and clean sound database
		\item structure correctly the disposition of the events
	\end{itemize}
\end{itemize}
\end{block}

\begin{block}{Proposition}
\begin{itemize}
	\item Use real recordings to base the distribution of the events
	\item Validate the realism of the sound mixtures obtained with a perceptual test
\end{itemize}

\end{block}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Table of contents}
  \tableofcontents
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{SimScene: a simulation tool}

\begin{frame}{\textit{SimScene}: a web-simulator to design sound environments (REF)}
\begin{block}{Principe}
\begin{itemize}
	\item Superpose sound events and a background from an isolated sound database
	\item Control a set of high level parameters
	\begin{itemize}
		\item presence time and sound level of each sound class
		\item number of occurrence of a sound class
	\end{itemize}
	\item Standard deviation on each parameter to bring random behavior
	\item Generate a separate audio file for each sound class present\\
	$\Rightarrow$ enable to know their exact contribution

\end{itemize}
\end{block}

\begin{block}{2 modes to design the sound mixtures}
\begin{itemize}
	\item \textit{abstract}: the user fills all the information 
	\item \textit{replicate}: the sound classes and the temporal position of each events come from an annotated text file
\end{itemize}
\end{block}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\textit{SimScene}: a simulation tool to design sound environments}

\begin{block}{Creation of the isolated sound database}
\begin{itemize}
	\item Sounds found online (\textit{fressound.org})
	\item Based on \textit{UrbanSound8k} database (REF)
\end{itemize}
\end{block}

\begin{block}{Recordings of car passages on Ifsttar-Nantes runway}
\begin{itemize}
	\item 4 cars (Renault Senic, Megane, Clio and Dacia Sandero)
	\item 104 car passages recorded for constant speeds (from 20 to 90 km/h), acceleration and braking phase for multiple gear ratios and stopped car 
\end{itemize}		
\end{block}

\begin{block}{To sum up}
\begin{itemize}
\item 154 samples of sound background (traffic, crowds, birds, rain \dots)
\item 245 samples of sound events (car passages, birds, car horns \dots)
\end{itemize}
\end{block}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Study of urban sound recordings}

\begin{frame}{Study of urban sound recordings}

\begin{block}{Corpus of urban recordings}
\begin{itemize}
	 \item Recordings made in the \nth{13} district in Paris (FR) (REF)
	 \item 19 recorded points on a define path
	 \item Recordings made on 2 days (03/23/2015 and 03/30/2015), twice a day
	 \item Among the 76 audio files available, 74 are used
	 \item Multiple sound environments 
	 \begin{itemize}
	 	\item \textit{Park}, \textit{quiet street}, \textit{noisy street} and \textit{very noisy street}
	 \end{itemize}
\end{itemize}
\end{block}

\begin{figure}[hbtp]
\centering
\includegraphics[width=.4\linewidth]{pictures/trajet_19pts.png}
\caption{Map of the soundwalk with the 19 stop points}
\end{figure}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Study of urban sound recordings}
\begin{itemize}
	\item Annotation of each audio file 
	\begin{itemize}
		\item sound classes, begin and end times
	\end{itemize}

\end{itemize}

\begin{block}{From the observations and the annotations}
\begin{itemize}
	\item Synthesis of the 74 real scenes with \textit{simScene} with the \textit{replicate} mode
	\item 15 seconds example of a simulated scene\\

\centering
\includemedia[addresource=15sExample_22.mp3, 
flashvars={source=15sExample_22.mp3&autoPlay=true}]{\fbox{Play}}{APlayer.swf}

\end{itemize}
\end{block}
\vspace{0.5cm}
$\Rightarrow$ Are the simulated scenes realistic enough ?

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Design of the perceptual test}

\begin{frame}{Design of the perceptual test}

\begin{block}{Objective}
\begin{itemize}
	\item Evaluation of the realism of the simulated scene among real recordings by a panel of listeners
\end{itemize}
\end{block}

\begin{itemize}
	\item 40 audio tested
	\begin{itemize}
		\item 30 seconds duration
		\item Normalization to the same the sound level
		\item 20 audio files from the real recordings and the same 20 replicated
	\end{itemize}

	\item 50 participants \\	
	$\Rightarrow$ Each judge cannot evaluate all the scene
\end{itemize}

\begin{block}{Partially Balanced Incomplete Block Design}
\begin{itemize}
	\item 20 audio files listen by each judge 
	\item Design the order of listening for each judge with a fair statistical distribution
	\item Each judge listen a mix of 10 real and 10 replicated scenes
\end{itemize}

\end{block}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Design of the perceptual test}
\begin{itemize}
	\item Web-page administered online on the 8 February 2017 and closed 12 days later
	\item For each scene listened
	\begin{itemize}
		\item answer to the question 'Is the scene you listen seems, to you, realistic ?'
		\item evaluation on a 7 points scale 
		\item possibility to re-listen the scene and to put a comment
	\end{itemize}
	\item The gender (H/F), the age and the experience in the listening of urban sound environments are then asked
\end{itemize}

\begin{block}{Constitution of the panel}
\begin{itemize}
	\item 18 females and 31 males (1 not documented)
	\item Average age : 36 ($\pm$ 12) years old
	\item 62 $\%$ of the participant declared having NO experience in the listening of urban sound mixtures
\end{itemize}
\end{block}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
%\begin{frame}{Results}
%
%\begin{block}{ANalyse Of VAriance}
%\begin{itemize}
%	\item ANOVA to evaluate 
%	\item $H_0$ hypothesis : the distribution of note between the real and the simulated scenes cannot be discernible
%	\item $H_0$ hypothesis rejected if \textit{p-value} $< \alpha = 0.05$ 
%\end{itemize}
%\end{block}
%
%\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Results}

\begin{block}{According to the 'type'}

\begin{figure}

\begin{flushleft}
\begin{minipage}[l]{.5\linewidth}
\includegraphics[width=\textwidth]{pictures/testPerceptif_boxplotType_EN.pdf}
\captionof{figure}{\footnotesize  Box-and-whiskers plot of the rating of realism according to the type of scene}
\end{minipage}
\hspace{0.1cm}
\begin{minipage}[r]{.4\linewidth}
\begin{itemize}
\footnotesize
	\item Hypothesis test $H_0$ : the distribution of the scores between the real and simulated scenes are similar
	\item Significance level $\alpha$ = 5 $\%$ to reject $H_0$
\end{itemize}

\footnotesize
\begin{tabular}{lccc}
	& DOF & $\vert t \vert$    & \textit{p-value} \\ 
\hline
type & 49 & 1.37 & 0.17    \\
\hline 
\end{tabular}
\captionof{table}{\footnotesize Paired sample \textit{t-test} }
\label{blablatab}
\end{minipage}

\end{flushleft}

\end{figure}
\end{block}

\begin{itemize}
\footnotesize
	\item T-test performed between the mean scores of the real and simulated scenes for each judge
	\item Real and simulated scenes evaluated in a similar way (\textit{p-value} $> \alpha = 0.05$)\\
\end{itemize}

$\Rightarrow$ Validation of the urban sound mixtures realism

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Results}

\begin{block}{According to the 'type', judge' and 'sound env.' factors}
\begin{figure}
\begin{flushleft}

\begin{minipage}[l]{.4\linewidth}		
\includegraphics[width=\textwidth]{pictures/testPerceptif_boxplotAmbianceCOLOR_EN.pdf}
\captionof{figure}{\footnotesize Data distribution for the 'scene type' and 'sound environment' factors}
\end{minipage}% à retirer pour voir la différence.	
\begin{minipage}[r]{.5\linewidth}
\footnotesize
\begin{tabular}{p{2.35cm}p{0.5cm}p{0.95cm}p{1.65cm}}
        &  DOF &  \textit{F-stat.}    &  \textit{p-value} \\ 
\hline
 type &  1 &  1.38 &  0.24   \\
\hline
 sound env.   &  3 &  4.69 &  3.60$\times 10^{-3}$  \\ 
\hline
 judge   &  49 &  5.00 &  0.18$\times 10^{-9}$  \\ 
\hline
 type/sound env. &  3 &  6.80 &  0.20$\times 10^{-3}$\\
\hline
 type/judge &  49 &   1.07 &   0.35\\
\hline
 sound env./judge &  147 &  1.43 &  1.60$\times 10^{-3}$\\
\end{tabular}
\captionof{table}{\footnotesize Three-ways ANOVA}
\label{tab:p_value_type_ambience}
\end{minipage}
\end{flushleft}
\end{figure}
\end{block}

\begin{itemize}
\footnotesize
	\item Distribution between real and simulated scenes non significant again
	\item Interaction between sound env./judge and type/sound env.\\
	$\Rightarrow$ Evolution of the distributions different according to the levels
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Results}
\begin{block}{Interaction between the 'type' and 'sound environment' factors}
\begin{minipage}[l]{0.5\linewidth}		
\begin{figure}[hbtp]
\centering
\includegraphics[width=\textwidth]{pictures/testPerceptif_interactionAmbianceCOLOR_EN.pdf}
\caption{Mean scores for the 4 sound environments according to the type}
\end{figure}
\end{minipage}% à retirer pour voir la différence.	
\begin{minipage}[c]{0.5\linewidth}
\begin{itemize}
\footnotesize
	\item We cannot conclude on the influence of the factors on the score
	\item Better mean scores for the 'noisy' and 'veryNoisy' simulated scenes
	\item Comments let by the panelists reveal that some noises are too loud (foot steps and birds) on the park and quiet simulated scenes
\end{itemize}

\end{minipage}
\end{block}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\begin{frame}{Conclusion}

\begin{itemize}
\item Design of a corpus of realistic urban scenes 
\item Based on the annotation of 74 real recordings 
\item Use of the simulation software \textit{simScene}
\item Representative sound database of isolated sounds with car passages recorded specifically
\item The realism of the simulated scene is validated by a perceptual test
	\begin{itemize}
	\item No notable difference between the evaluation of the real and the simulated scenes
	\item The 'park' and 'quiet street' atmospheres are both perceived as less realistic than the 'noisy' and 'very noisy street' atmosphere
	\item some improvement can be made on the sound level of some sound classes (bird, foot step)
	\end{itemize}
\item Urban sound scene corpus that can be use to test sound recognition, classification or separation tools or to realize specific perceptual tests
\end{itemize}

\end{frame}

\begin{frame}

\titlepage

\end{frame}

\end{document}

### 2017-06-19 boston-presentation.tex
fex fixes

@@
 \author[J. Gloaguen]{\emph{Jean-R\'emy Gloaguen}$^1$,  Arnaud Can$^1$,\\ Mathieu Lagrange$^2$,  Jean-Fran\c cois Petiot$^2$}
-\institute[Ifsttar]{$^1$French Institute of Science and Technology for Transport, Development and Networks, \\ Environmental Acoustics Laboratory \\Allée des Ponts et Chaussées, Route de Bouaye, - CS4, 44344 Bouguenais Cedex, FR  \\ \vspace{.25cm} $^2$Nantes' Numeric Sciences Laboratory, \\1, rue de la Noë, BP 92101, 44321 Nantes Cedex 3, FR}
+\institute[Ifsttar]{$^1$French Institute of Science and Technology for Transport, Development and Networks, \\ Environmental Acoustics Laboratory \\Allée des Ponts et Chaussées, Route de Bouaye, - CS4, 44344 Bouguenais Cedex, FR  \\ \vspace{.25cm} $^2$ LS2N, \\1, rue de la Noë, BP 92101, 44321 Nantes Cedex 3, FR}
 \date[06/29/2017]{Acoustics' 17 Boston\\06/29/2017}
@@
 	\item Based on real sound environments with \textit{soundwalks} or recordings listened on laboratory\\
-	$\Rightarrow$ ecological validity \\
+	$\Rightarrow$ good ecological validity \\
 	$\Rightarrow$ no controlled framework \\
@@
 		$\Rightarrow$ creation of specific urban sound environments \\
-		$\Rightarrow$ no ecological validity \\
+		$\Rightarrow$ low ecological validity \\
 \end{itemize}
@@
 \begin{itemize}
-	\item Use real recordings to base the distribution of the events
-	\item Validate the realism of the sound mixtures obtained with a perceptual test
+	\item Transcribe real recordings
+	\item use the transcription to simulate controled stimuli
+	\item Validate the realism of the simulated sound mixtures obtained with a perceptual test
 \end{itemize}
@@
 
-\begin{frame}{\textit{SimScene}: a web-simulator to design sound environments (REF)}
+\begin{frame}{\textit{SimScene}: a simulator to design sound environments \footnote{\url{https://bitbucket.org/mlagrange/simscene}}}
 \begin{block}{Principe}
 \begin{itemize}
-	\item Superpose sound events and a background from an isolated sound database
+	\item Mix sound events and a background taken from an isolated sound database
 	\item Control a set of high level parameters
 	\begin{itemize}
-		\item presence time and sound level of each sound class
+		\item onset time and sound level of each sound class
 		\item number of occurrence of a sound class
@@
 	\item Standard deviation on each parameter to bring random behavior
-	\item Generate a separate audio file for each sound class present\\
+	\item Generate a separate audio file for each sound class \\
 	$\Rightarrow$ enable to know their exact contribution
@@
 \begin{itemize}
-	\item \textit{abstract}: the user fills all the information 
+	\item \textit{abstract}: the user sets abstract information (mean / variance)
 	\item \textit{replicate}: the sound classes and the temporal position of each events come from an annotated text file
@@
 
-\begin{block}{Constitution of the panel}
+\begin{block}{The panel}
 \begin{itemize}
